### 2022-02-28 paper/paper.tex
Moved code into new directory and added plot_results.py (2nd attempt).

\documentclass[a4paper,abstracton]{scrartcl}

% author and affiliation block
\usepackage{authblk}
\renewcommand\Affilfont{\small}

% maths packages
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{bm}

% bibliography and referencing
\usepackage{natbib}
\bibliographystyle{apalike}
\usepackage{hyperref}
\usepackage{xcolor}
\definecolor{darkgreen}{RGB}{0,127,173}
\definecolor{darkblue}{RGB}{0,0,200}
\hypersetup{colorlinks, linkcolor={darkblue}, citecolor={darkgreen}, urlcolor={darkblue}}

% figures and graphics
\usepackage{graphicx}
\graphicspath{{figures/}}  % path to directory containing all figures

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%  Macros
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\F}{{\mathcal{F}}}
\newcommand{\M}{{\mathcal{M}}}
\newcommand{\U}{{\mathcal{U}}}
\newcommand{\X}{{\mathcal{X}}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{A faster exact method for solving the robust multi-mode resource-constrained project scheduling problem}

\author[1]{Matthew Bold\footnote{Corresponding author, email: m.bold1@lancaster.ac.uk}}
\author[2]{Marc Goerigk}

\affil[1]{STOR-i Centre for Doctoral Training, Lancaster University, Lancaster, UK}	
\affil[2]{Network and Data Science Management, University of Siegen, Siegen, Germany}

\date{}

\maketitle

\begin{abstract}
\end{abstract}

\noindent\textbf{Keywords:} project scheduling; optimisation under uncertainty; robust optimisation; budgeted uncertainty

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
The multi-mode resource-constrained project scheduling problem (MRCPSP) is a generalisation of the widely-studied resource-constrained project scheduling problem (RCPSP) to include multiple processing modes for each activity. The purpose of introducing activity processing modes is to model situations in which there is more than one way of executing project activities, with each option having its own duration and resource requirements. The MRCPSP consists selecting the processing modes and start times for a given set of activities, subject to a set of precedence constraints and limited resource availability, with the objective of minimising the overall project duration, known as the makespan.

The work in the paper extends the compact reformulation for the RCPSP developed by \cite{bold2021compact} for application to the MRCPSP. A computational study is presented comparing this approach against the Benders' decomposition approach used in \cite{balouka2021robust} for solving MRCPSP instances with uncertain activity durations. Note that the method used in \cite{balouka2021robust} is an extension of the approach used by \cite{bruni2017adjustable} to solve the RCPSP with uncertain activity durations.

\section{Problem description}

A project consists of a set of non-preemptive activities $V=\{0,1,\dots,n,n+1\}$, where 0 and $n+1$ denote the dummy source and sink activities respectively. Each activity $i\in V$ has a set of available processing modes given by $M_i=\{1,\dots,|M_i|\}$. We denote a particular choice of mode selections as $\bm{m}=(m_1,\dots,m_n)$, and the set of all possible mode selections by $\M\subseteq \mathbb{N}^n$. The nominal duration of activity $i$ when executed in mode $m_i\in M_i$ is given by $\hat{d}_{im_i}$, whilst its worst-case duration is given by $\bar{d}_{im_i}+\hat{d}_{im_i}$, where $\hat{d}_{im_i}$ is its maximum durational deviation. Each mode for activity $i$, $m_i\in M_i$, has an associated renewable resource requirement of $r_{im_ik}$ for each $k\in K$, where $K$ is the set of renewable resource types involved in the project. Each renewable resource $k\in K$ has an availability of $R^_{k}$ at each time period in the project horizon. As well as resource constraints, the project is subject to a set of strict finish-to-start precedence constraints given by $E$, where $(i,j)\in E$ enforces that activity $i$ must finish before activity $j$ can begin. A project can be represented on a directed graph $G(V,E)$. /FIGURE shows an example project involving seven non-dummy activities and a single renewable resource.

For a given choice of processing modes for each activity $\bm{m}$, we assume that the activity durations lie somewhere in a budgeted uncertainty set of the form
$$\U_{\bm{m}}(\Gamma)=\Bigg\{\bm{d}\in\mathbb{R}_+^{|V|}:d_{im_i}=\bar{d}_{im_i}+\zeta_i\hat{d}_{im_i},\,0\leq \zeta_i \leq 1,\,\forall i\in V,\,\sum_{i\in V}\zeta_i \leq \Gamma\Bigg\}.$$
Introduced by \cite{bertsimas2004price}, the motivation of the budgeted uncertainty set is to control the pessimism of the solution by introducing a robustness parameter $\Gamma$ to limit the number of jobs that can simultaneously achieve their worst-case durations. Observe that when $\Gamma=0$, each activity takes its nominal duration and the resulting problem is the deterministic MRCPSP. At the other extreme, when $\Gamma=n$, every activity takes its worst-case duration, in which case the uncertainty set becomes equivalent to an interval uncertainty set. In this case the problem can again be solved as a deterministic MRCPSP instance considering only worst-case durations.

For a given selection of activity modes $\bm{m}$, a forbidden set is defined to be any subset of activities $F_{\bm{m}}\subseteq V$ that are not precedence-related, such that $\sum_{i\in F_{\bm{m}}}r_{im_ik}>R_k$ for at least one resource $k\in K$. That is, a forbidden set is a collection of activities that cannot be executed in parallel only because of resource limitations.

Applying the main representation theorem of \cite{bartusch1988scheduling}, given a particular choice of activity modes $\bm{m}$, a solution to the MRCPSP can be defined by a set of additional precedences $X_{\bm{m}}\subseteq V^2\setminus E$ such that the extended precedence network $G(V,E\cup X_{\bm{m}})$ is acyclic and contains no forbidden sets. Such an extension to the project network is referred to as a \textit{sufficient selection}.

Hence the two-stage robust MRCPSP under budgeted uncertainty, which we refer to as the robust MRCPSP, can be written as
\begin{align}
	& \min_{\bm{m}\in \M,\,X_{\bm{m}}\in \X_{\bm{m}}} \max_{\bm{d}\in\U_{\bm{m}}(\Gamma)} \min_{\bm{S}} S_{n+1}\label{eqn:robust_mrcpsp_1}\\
	& S_0 = 0\\
	& S_j - S_i \geq d_{im_i} \qquad \forall (i,j)\in E\cup X_{\bm{m}}\\
	& S_i \geq 0 \qquad \forall i\in V.\label{eqn:robust_mrcpsp_4}
\end{align}
This problem aims to selects activity modes and a corresponding sufficient selection in order to minimise the worst-case project makespan.

\section{A compact reformulation}

We propose solving this problem using an extended version of the compact reformulation from \cite{bold2021compact}. The development of this formulation for the RCPSP involved a reformulation of the adversarial sub-problem of determining the worst-case distribution of $\Gamma$ delays among the project activities. The reformulated sub-problem took the form of a longest-path problem through an augmented project network.

The derivation for the equivalent compact reformulation for the MRCPSP is more or less identical to the one presented in that paper for the RCPSP. We therefore omit its derivation here, and instead just present the resulting model.

\begin{align}
\min \ & S_{n+1,\Gamma}\label{eqn:compact1}\\
\text{s.t. } & S_{00}=0\label{eqn:compact2}\\
	     & S_{j\gamma} - S_{i\gamma} \geq \bar{d}_{im}x_{im} - N(1-y_{ij}) & \hspace{-5cm}\forall (i,j)\in V^2,\,\forall m\in M_i,\,\gamma=0,\dots,\Gamma\label{eqn:compact_bigM1}\\
	     & S_{j,\gamma+1} - S_{i\gamma} \geq (\bar{d}_{im} + \hat{d}_{im})x_{im} - N(1-y_{ij})\nonumber \\
	     &&\hspace{-5cm} \forall (i,j)\in V^2,\, m\in M_i,\,\gamma = 0,\dots,\Gamma-1\label{eqn:compact_bigM2}\\
	     & y_{ij}=1 & \forall (i,j) \in E\cup\{(n+1,n+1)\}\label{eqn:compact5}\\
	     & f_{ijk}\leq P_{ijk}y_{ij} & \hspace{-5cm}\forall (i,j)\in V^2,\,i\neq n+1,\,j\neq0,\, \forall k \in K\label{eqn:compact6}\\
	     & \sum_{i\in V\setminus\{n+1\}}f_{ijk}=\sum_{m\in M_j}r_{jmk}x_{jm} & \forall j \in V\setminus\{0\} ,\, \forall k\in K\label{eqn:compact7}\\
	     & \sum_{j\in V\setminus\{0\}}f_{ijk}=\sum_{m\in M_i}r_{imk}x_{im} & \forall i \in V\setminus\{n+1\},\,\forall k\in K\label{eqn:compact8}\\
	     & \sum_{m\in M_i}x_{im} = 1 & \forall i\in V\label{eqn:compact9}\\
	     & \sum_{m\in M_i}r^{\nu}_{imk}x_{im} \leq R^{\nu}_k & \forall i \in V,\,k\in K^{\nu}\\
	     & S_{i\gamma}\geq 0 &\forall i\in V,\,\gamma \in 0,\dots, \Gamma\label{eqn:compact10}\\
	     & f_{ijk}\geq 0 & \forall (i,j)\in V^2 ,\, \forall k \in K\label{eqn:compact11}\\
	     & y_{ij}\in\{0,1\} & \forall (i,j)\in V^2\label{eqn:compact12}\\
	     & x_{im}\in\{0,1\} & \forall i\in V,\,m\in M_i,
\end{align}
where $N=\sum_{i\in V}\max_{m\in M_i}(\bar{d}_{im}+\hat{d}_{im})$ is an upper bound on the minimum makespan, and $P_{ijk}=\min\{\max_{m\in M_i}r_{imk},\,\max_{m\in M_j} r_{jmk}\}$ is the maximum possible flow of resource $k$ from $i$ into $j$.

This formulation can be solved directly by a standard mathematical optimisation solver such as Gurobi or CPLEX.

\section{A Benders' decomposition approach}

In this section we outline the Benders' decomposition approach to solving the robust MRCPSP first presented in \cite{balouka2021robust}.

\section{Computational experiments and results}

\section{Conclusions}

\section*{Acknowledgements}

The authors are grateful for the support of the EPSRC-funded (EP/L015692/1) STOR-i Centre for Doctoral Training.

\bibliography{paper}

\end{document}

### 2022-02-28 code/benders.py

@@ -1,0 +1,310 @@
+import time
+import gurobipy as gp
+from gurobipy import GRB
+
+
+class Benders:
+    """
+    Class to solve uncertain MRCPSP instance using Bender' style decomposition.
+    """
+
+    def __init__(self, instance, Gamma, time_limit):
+        """
+        Initialises Benders object with instance to solve and algorithm parameters.
+
+        :param instance: Robust MRCPSP instance to solve
+        :type instance: Instance
+        :param Gamma: Robustness parameter, i.e. max number of jobs that can achieve worst-case durations
+            simultaneously. Takes an integer value from 0 to n (= number of non-dummy jobs in instance).
+        :type Gamma: int
+        :param time_limit: Max running time of algorithm in seconds. Current UB is returned if time_limit is reached.
+        :type time_limit: int
+        """
+        self.instance = instance
+        self.Gamma = Gamma
+        self.time_limit = time_limit
+
+        self.LB = float('-inf')  # LB on objective function of robust MRCPSP instance
+        self.UB = float('inf')  # UB on objective function of robust MRCPSP instance
+        self.cuts = []  # cuts applied to master problem
+
+        # Master problem model and variables. Populated in create_master_problem and updated when solved.
+        self.master_model = None
+        self.master_eta = None
+        self.master_y = None
+        self.master_x = None
+        self.master_f = None
+        self.master_s = None
+        self.master_phi = None
+        # Sub-problem model, objective value and parameters. Updated each time sub-problem is generated and solved.
+        self.sub_model = None
+        self.sub_objval = None
+        self.sub_alpha = None
+        self.sub_w = None
+        self.sub_xi = None
+
+    def solve(self, num_threads=4, print_log=False):
+        """
+        Runs Benders' decomposition algorithm to solve robust MRCPSP instance. Terminates upon finding an optimal
+        solution or reaching the specified time limit.
+
+        :param num_threads: Number of threads to use when solving. Defaults to 4.
+        :type num_threads: int
+        :param print_log: Indicates whether or not to print Gurobi solve log to terminal. Defaults to False.
+        :type print_log: bool
+        :return: Dictionary containing solution information.
+        :rtype: dict
+        """
+        # Benders' decomposition algorithm
+        t = 1  # iteration number
+        iteration_times = []  # List to store time of each iteration. Average iteration time is reported in solution.
[…]

### 2022-02-28 code/compact_reformulation.py

@@ -1,0 +1,115 @@
+import gurobipy as gp
+from gurobipy import GRB
+
+
+class CompactRefomulation:
+    """
+    Class to solve uncertain MRCPSP instance using compact reformulation.
+    """
+
+    def __init__(self, instance, Gamma, time_limit):
+        """
+        Initialises CompactReformulation object with instance to solve and solve parameters.
+
+        :param instance: Robust MRCPSP instance to solve
+        :type instance: Instance
+        :param Gamma: Robustness parameter, i.e. max number of jobs that can achieve worst-case durations
+                simultaneously. Takes an integer value from 0 to n.
+        :type Gamma: int
+        :param time_limit: Time limit for Gurobi, in seconds
+        :type time_limit: int
+        :return: Dictionary containing solution information.
+        :rtype: dict
+        """
+        self.instance = instance
+        self.Gamma = Gamma
+        self.time_limit = time_limit
+
+    def solve(self, num_threads=4, print_log=False):
+        """
+        Builds and solves compact reformulation for uncertain MRCPSP. Terminates upon finding an optimal solution or
+        reaching the specified time limit.
+
+        :param num_threads: Number of threads to use when solving. Defaults to 4.
+        :type num_threads: int
+        :param print_log: Indicates whether or not to print Gurobi solve log to terminal. Defaults to False.
+        :type print_log: bool
+        :return: Dictionary containing solution information.
+        :rtype: dict
+        """
+        # get instance data
+        V = self.instance.V
+        n = self.instance.n
+        E = self.instance.E
+        d = [self.instance.jobs[j].d for j in V]  # d[j][m] = duration of j in mode m
+        d_bar = [self.instance.jobs[j].d_bar for j in V]  # d_bar[j][m] = max durational deviation of j in mode m
+        M = [self.instance.jobs[j].M for j in V]  # M[j] = list of modes of j
+        # r[j][m][k] = amount of resource k required by j in mode m. Dummy source and sink all available resource.
+        r = [[self.instance.R]] + [self.instance.jobs[j].r for j in self.instance.N] + [[self.instance.R]]
+
+        # Create model
+        model_name = '{}_compact_reformulation'.format(self.instance.name)
[…]

### 2022-02-28 code/experiment.py

@@ -1,0 +1,108 @@
+import os
+import re
+from mrcpsp import load_nominal_mrcpsp
+from benders import Benders
+from compact_reformulation import CompactRefomulation
+
+
+def create_instances(instance_dir, uncertainty_level):
+    """
+    Returns list of uncertain MRCPSP instances generated from nominal MRCPSP instances contained in instance_dir. Max
+    durational deviations are set according to uncertainty_level.
+
+    :param instance_dir: Absolute filepath of directory containing nominal MRCPSP instances downloaded from PSPLIB.
+    :type instance_dir: str
+    :param uncertainty_level: Non-negative value representing desired ratio between d_bar[m] / d[m].
+    :type uncertainty_level: float
+    :return: List of uncertain MRCPSP instances.
+    :rtype: list
+    """
+    instance_files = [instance for instance in os.listdir(instance_dir)]
+    # sort instance files into their natural order
+    instance_files = sorted(instance_files, key=lambda f: 10 * int(''.join(re.findall(r'\d+', f[3:5]))) + int(
+        ''.join(re.findall(r'\d+', f[5:8]))))
+
+    instances = []
+    for file in instance_files:
+        # load nominal instance from data file
+        instance_filepath = os.path.join(instance_dir, file)
+        instance = load_nominal_mrcpsp(instance_filepath)
+        # set durational uncertainty
+        instance.set_dbar_uncertainty_level(uncertainty_level)
+        # append to list of uncertain MRCPSP instances
+        instances.append(instance)
+    return instances
+
+
+def benders_experiment(instances, Gamma, time_limit):
+    """
+    Solves set uncertain MRCPSP instances using Bender' decomposition and writes results to an output file.
+
+    :param instances: List of uncertain MRCPSP instances to solve.
+    :type instances: list
+    :param Gamma: Robustness parameter to use.
+    :type Gamma: int
+    :param time_limit: Time limit in seconds to give to Benders' algorithm for each instance.
+    :type time_limit: int
+    """
+    # create results file
+    current_dir = os.path.dirname(os.path.realpath(__file__))
+    results_file = os.path.join(current_dir, 'benders_results_{}.txt'.format(Gamma))
+    f = open(results_file, 'w+')
+    f.write("instance\tGamma\tobjval\tobjbound\tgap\tn_iterations\tavg_iteration\truntime\n")
+    f.close()
+
+    # solve each instance using Benders' and write results
+    for instance in instances:
[…]

### 2022-02-28 code/mrcpsp.py

@@ -1,0 +1,219 @@
+import re
+import os
+import math
+
+
+def load_nominal_mrcpsp(filepath):
+    """
+    Loads MRCPSP instance from data file and returns Instance object.
+
+    :param filepath: Filepath of MRCPSP instance to load
+    :type filepath: str
+    :return: MRCPSP instance
+    :rtype: Instance
+    """
+    with open(filepath, 'r') as f:
+        raw_lines = f.read().splitlines()
+
+    block = 0
+    succ = {}  # dict of successors of each job
+    M = {}  # dict of modes for each job
+    d = {}  # dict of duration for each job
+    r = {}  # dict of resource requirements for each job
+    for line in raw_lines:
+        ints = list(map(int, re.findall(r'\d+', line)))  # list of integers in line
+        if len(ints) == 0:
+            if '*****' in line:
+                block += 1
+            continue
+        elif block == 2:
+            if 'jobs' in line:
+                n = int(ints[0]) - 2  # number of jobs excl. dummy source and sink
+            elif 'horizon' in line:
+                T = int(ints[0])  # project horizon
+            elif '- renewable' in line:
+                n_renew = int(ints[0])  # number of renewable resources
+            elif '- nonrenewable' in line:
+                n_nonrenew = int(ints[0])  # number of non-renewable resources
+        elif block == 4:
+            jobnr = int(ints[0]) - 1  # jobs are numbered from 1 to n+2 in data files
+            M[jobnr] = [m for m in range(int(ints[1]))]
+            succ[jobnr] = [int(succ) - 1 for succ in ints[3:]]  # ints[3:] = successors
+        elif block == 5:
+            n_res = n_renew + n_nonrenew
+            if len(ints) == 3 + n_res:  # mode 1 data
+                jobnr = int(ints[0]) - 1
+                d[jobnr] = [int(ints[2])]
+                r[jobnr] = [[int(x) for x in ints[3:]]]
+            elif len(ints) == 2 + n_res:  # modes 2,3,4... data
+                d[jobnr].append(int(ints[1]))
+                r[jobnr].append([int(x) for x in ints[2:]])
+        if block == 6:
+            R = [int(x) for x in ints]
+    jobs = {}
+    for j in range(n + 2):
+        pred = [i for i in range(j) if j in succ[i]]
+        jobs[j] = NominalJob(j, pred, succ[j], M[j], d[j], r[j])
[…]

### 2022-02-28 code/parallel_experiment.py

@@ -1,0 +1,238 @@
+import os
+import re
+import argparse
+import multiprocessing
+from filelock import FileLock
+from functools import partial
+from mrcpsp import load_nominal_mrcpsp
+from benders import Benders
+from compact_reformulation import CompactRefomulation
+
+
+def create_instances(instance_dir, uncertainty_level):
+    """
+    Returns list of uncertain MRCPSP instances generated from nominal MRCPSP instances contained in instance_dir. Max
+    durational deviations are set according to uncertainty_level.
+
+    :param instance_dir: Absolute filepath of directory containing nominal MRCPSP instances downloaded from PSPLIB.
+    :type instance_dir: str
+    :param uncertainty_level: Non-negative value representing desired ratio between d_bar[m] / d[m].
+    :type uncertainty_level: float
+    :return: List of uncertain MRCPSP instances.
+    :rtype: list
+    """
+    instance_files = [instance for instance in os.listdir(instance_dir)]
+    # sort instance files into their natural order
+    instance_files = sorted(instance_files, key=lambda f: 10 * int(''.join(re.findall(r'\d+', f[3:5]))) + int(
+        ''.join(re.findall(r'\d+', f[5:8]))))
+
+    instances = []
+    for file in instance_files:
+        # load nominal instance from data file
+        instance_filepath = os.path.join(instance_dir, file)
+        instance = load_nominal_mrcpsp(instance_filepath)
+        # set durational uncertainty
+        instance.set_dbar_uncertainty_level(uncertainty_level)
+        # append to list of uncertain MRCPSP instances
+        instances.append(instance)
+    return instances
+
+
+def benders_solve_and_write(instance, Gamma, time_limit, num_threads):
+    """
+    Solves uncertain MRCPSP instance using Bender' decomposition and writes result to an (existing) output file.
+
+    :param instance: Uncertain MRCPSP instance to solve.
+    :type instance: Instance
+    :param Gamma: Robustness parameter to use.
+    :type Gamma: int
+    :param time_limit: Time limit in seconds to give to Benders' algorithm for each instance.
+    :type time_limit: int
+    :param num_threads: Number of threads to use when solving instance.
+    :type num_threads: int
+    """
+    current_dir = os.path.dirname(os.path.realpath(__file__))
+    results_file = os.path.join(current_dir, 'benders_results_{}.txt'.format(Gamma))
+
+    # solve instance using Benders' and write result
[…]

### 2022-02-28 code/plot_results.py

@@ -1,0 +1,21 @@
+
+
+def read_benders_results(results_file):
+    """
+    Reads text file of Benders' results from results_file and returns dictionary of results data.
+
+    :param results_file: Absolute filepath to results file produced by running a Benders' experiment.
+    :type results_file: str
+    :return: Dictionary of results data.
+    :rtype: dict
+    """
+    print(results_file)
+    results = {}
+    with open(results_file, 'r') as f:
+        raw_lines = f.read().splitlines()
+
+    for line in raw_lines[1:]:
+        splitline = line.split()
+        print(raw_lines)
+
+read_benders_results('/home/boldm1/OneDrive/robust-mrcpsp/code/results/storm_j20_7200/benders_results_0.txt')

### 2022-02-28 code/solve_instance.py

@@ -1,0 +1,40 @@
+from benders import Benders
+from compact_reformulation import CompactRefomulation
+
+
+def solve(instance, solve_method, Gamma, time_limit):
+    """
+    Solves given uncertain MRCPSP instance using solve_method ('compact_reformulation' or 'benders') and prints solve
+    log to terminal. Used to test code for specific instances.
+
+    :param instance: Uncertain MRCPSP instance.
+    :type instance: Instance
+    :param solve_method: Solution method to use, i.e. 'compact_reformulation' or 'benders'.
+    :type solve_method: str
+    :param Gamma: Robustness parameter, i.e. max number of jobs that can achieve worst-case durations simultaneously.
+        Takes integer value from 0 to n (= number of non-dummy jobs in instance).
+    :type Gamma: int
+    :param time_limit: Time limit (in seconds) to give to chosen solution method.
+    :type time_limit: int
+    """
+    # check solve_method is recognised
+    if solve_method not in ['compact_reformulation', 'benders']:
+        raise Exception("{} is not a recognised solve method. Available methods are 'compact_reformulation' and "
+                        "'benders'. Please check this input and try again.".format(solve_method))
+
+    # solve instance using compact reformulation
+    elif solve_method == 'compact_reformulation':
+        print("Solving {} using compact reformulation:".format(instance.name))
+        print('"""""""""""""""""""""""""""""""""""""""""""""\n')
+        sol = CompactRefomulation(instance, Gamma, time_limit).solve(print_log=True)
+
+    # solve instance using Benders' decomposition
+    elif solve_method == 'benders':
+        print("Solving {} using Benders' decomposition:".format(instance.name))
+        print('"""""""""""""""""""""""""""""""""""""""""""""\n')
+        sol = Benders(instance, Gamma, time_limit).solve(print_log=True)
+
+    print(sol)
+    print("objval:", sol['objval'])
+    print("objbound:", sol['objbound'])
+    print("runtime:", sol['runtime'])

### 2022-02-28 code/submit.sh

@@ -1,0 +1,22 @@
+#PBS -N experiment  # job name
+#PBS -o output/  # output path
+#PBS -j oe  # merge error messages with output
+#PBS -l ncpus=20  # limit total number of cpus
+
+export num_threads=4 # number of threads to use for each process
+export num_processes=5 # total number of processes to use. num_treads * num_processes = ncpus.
+
+# Set Gurobi environment variables
+export GUROBI_HOME="/opt/gurobi901/linux64"
+export PATH="${PATH}:${GUROBI_HOME}/bin"
+export LD_LIBRARY_PATH="${LD_LIBRARY_PATH}:${GUROBI_HOME}/lib"
+
+# Copy and paste this block into all STORM submission scripts. Limits thread usage of various packages by num_threads.
+export MKL_NUM_THREADS=$num_threads      # Limits Intel math kernel
+export OPENBLAS_NUM_THREADS=$num_threads # Limits OPENBLAS, this is the most important one
+export MC_CORES=$num_threads             # Limits some packages in R
+export OMP_NUM_THREADS=$num_threads      # Limits OpenMP
+export NUMEXPR_NUM_THREADS=$num_threads  # Limits NumExpr in python
+
+cd ~/robust-mrcpsp
+python parallel_experiment.py -instance_dir=/home/boldm1/robust-mrcpsp/instances/j10.mm -solve_method=compact_reformulation -Gamma=0 -time_limit=3600 -num_threads=$num_threads -num_processes=$num_processes

### 2022-02-28 code/toy_example1.py

@@ -1,0 +1,43 @@
+
+import mrcpsp
+from compact_reformulation import CompactRefomulation
+from benders import Benders
+
+"""
+Testing code by solving toy example from Balouka and Cohen (2021).
+"""
+
+# read nominal data
+file_path = '/home/boldm1/OneDrive/robust-mrcpsp/instances/mrcpsp_toy_example_data.txt'
+instance = mrcpsp.load_nominal_mrcpsp(file_path)
+# set d_bar values explicitly
+instance.set_dbar_explicitly({0: [0], 1: [0, 3], 2: [8, 2], 3: [1], 4: [3], 5: [0], 6: [0]})
+
+# solve parameters
+Gamma = 3
+time_limit = 20
+
+print(f"\nSolving toy example given in {file_path}...")
+print("\nJob data [r, d, d_bar]:")
+print("-------------------------")
+for i in instance.V:
+    print("job {}:".format(i),
+          [[instance.jobs[i].r[m][0], instance.jobs[i].d[m], instance.jobs[i].d_bar[m]] for m in instance.jobs[i].M])
+
+print("\nCompact reformulation")
+print("---------------------")
+compact_sol = CompactRefomulation(instance, Gamma, time_limit).solve(print_log=False)
+print("objval:", compact_sol['objval'])
+print("runtime:", compact_sol['runtime'])
+print("modes:", compact_sol['modes'])
+print("network:", compact_sol['network'])
+print("resource flows:", compact_sol['flows'])
+
+print("\nBenders'")
+print("--------")
+benders_sol = Benders(instance, Gamma, time_limit).solve(print_log=False)
+print("objval:", benders_sol['objval'])
+print("runtime:", benders_sol['runtime'])
+print("modes:", benders_sol['modes'])
+print("network:", compact_sol['network'])
+print("resource flows:", benders_sol['flows'])

### 2022-03-01 code/plot_results.py
Plotting gaps and performance profile.

@@ -1,21 +1,194 @@
+import os
+import numpy as np
+import matplotlib.pyplot as plt
 
 
-def read_benders_results(results_file):
+def read_results(results_dir, solve_methods, Gammas):
     """
-    Reads text file of Benders' results from results_file and returns dictionary of results data.
+    Reads specified results files from directory and returns dictionary of results data.
 
-    :param results_file: Absolute filepath to results file produced by running a Benders' experiment.
-    :type results_file: str
-    :return: Dictionary of results data.
+    :param results_dir: Absolute filepath of directory containing results files. Directory must contain results files
+        for all combinations of solve methods and values of Gamma.
+    :type results_dir: str
+    :param solve_methods: List of solution methods to read results for.
+    :type solve_methods: list
+    :param Gammas: List of Gamma values to read results for.
+    :type Gammas: list
+    :return: Dictionary of results for each combination of solve method and Gamma value.
     :rtype: dict
     """
-    print(results_file)
-    results = {}
-    with open(results_file, 'r') as f:
-        raw_lines = f.read().splitlines()
+    # check results files exist for each combination solve_method and Gamma
+    missing_files = []
+    results_files = [f for f in os.listdir(results_dir) if os.path.splitext(f)[1] == '.txt']
+    for method in solve_methods:
+        for g in Gammas:
+            required_file = f"{method}_results_{g}.txt"
+            if required_file not in results_files:
+                missing_files.append(required_file)
+    if len(missing_files) > 0:
+        print("The following results files have not been found:")
+        for v in missing_files:
+            print(f"\t{v}")
+        raise Exception("Please add the missing files to the specified results directory and try again.")
 
-    for line in raw_lines[1:]:
-        splitline = line.split()
-        print(raw_lines)
+    # get data from results files
+    results_data = {}
+    for method in solve_methods:
+        results_data[method] = {}
+        for g in Gammas:
+            file_name = f"{method}_results_{g}.txt"
+            file_path = os.path.join(results_dir, file_name)
+            file_data = {}
+            with open(file_path, 'r') as f:
+                raw_lines = f.read().splitlines()
+                for line in raw_lines[1:]:
+                    splitline = line.split()
+                    instance = splitline[0]
+                    if method == 'benders':
[…]

### 2022-03-02 code/benders.py
Reducing number of variables in Benders subproblem.

@@ -235,10 +235,13 @@ class Benders:
         V = self.instance.V
         M = [self.instance.jobs[j].M for j in V]
 
+        # transitive network from master problem
+        T_E = [(i, j) for i in V for j in V if i != j if self.master_y[i, j].X > 0.99]
+
         # get mode selection from master problem
         master_M = [int(sum(m * self.master_x[i, m].X for m in M[i])) for i in V]
         # get longest path from sub-problem
-        sub_path = [(i, j) for i in V for j in V if self.sub_alpha[i, j].X > 0.99]
+        sub_path = [(e[0], e[1]) for e in T_E if self.sub_alpha[e[0], e[1]].X > 0.99]
 
         # some number larger than number of extra precedences in optimal master solution
         N = self.instance.n * self.instance.n
@@ -285,8 +288,8 @@ class Benders:
         model.setParam('Threads', num_threads)
 
         # Add variables to model
-        alpha = model.addVars([(i, j) for i in V for j in V], name="alpha", vtype=GRB.BINARY)
-        w = model.addVars([(i, j) for i in V for j in V], name="w", vtype=GRB.CONTINUOUS, lb=0)
+        alpha = model.addVars([(e[0], e[1]) for e in T_E], name="alpha", vtype=GRB.BINARY)
+        w = model.addVars([(e[0], e[1]) for e in T_E], name="w", vtype=GRB.CONTINUOUS, lb=0)
         xi = model.addVars([i for i in V], name="xi", vtype=GRB.CONTINUOUS, lb=0, ub=1)
 
         # Add constraints to model

### 2022-03-02 code/experiment.py

@@ -48,22 +48,21 @@ def benders_experiment(instances, Gamma, time_limit):
     # create results file
     current_dir = os.path.dirname(os.path.realpath(__file__))
     results_file = os.path.join(current_dir, 'benders_results_{}.txt'.format(Gamma))
-    f = open(results_file, 'w+')
-    f.write("instance\tGamma\tobjval\tobjbound\tgap\tn_iterations\tavg_iteration\truntime\n")
-    f.close()
+    with open(results_file, 'w+') as f:
+        f.write("instance\tGamma\tobjval\tobjbound\tgap\tn_iterations\tavg_iteration\truntime\n")
 
     # solve each instance using Benders' and write results
     for instance in instances:
         benders_sol = Benders(instance, Gamma, time_limit).solve()
         gap = abs(benders_sol['objbound'] - benders_sol['objval']) / abs(benders_sol['objval'])  # optimality gap
-        f = open(results_file, 'a')
-        f.write(
-            "{}\t{}\t{:.3f}\t{:.3f}\t{:.3f}\t{}\t{:.3f}\t{:.3f}\n".format(instance.name, Gamma, benders_sol['objval'],
-                                                                          benders_sol['objbound'], gap,
-                                                                          benders_sol['n_iterations'],
-                                                                          benders_sol['avg_iteration'],
-                                                                          benders_sol['runtime']))
-        f.close()
+        with open(results_file, 'a') as f:
+            f.write(
+                "{}\t{}\t{:.3f}\t{:.3f}\t{:.3f}\t{}\t{:.3f}\t{:.3f}\n".format(instance.name, Gamma,
+                                                                              benders_sol['objval'],
+                                                                              benders_sol['objbound'], gap,
+                                                                              benders_sol['n_iterations'],
+                                                                              benders_sol['avg_iteration'],
+                                                                              benders_sol['runtime']))
 
 
 def compact_reformulation_experiment(instances, Gamma, time_limit):
@@ -80,29 +79,28 @@ def compact_reformulation_experiment(instances, Gamma, time_limit):
     # create results file
     current_dir = os.path.dirname(os.path.realpath(__file__))
[…]

### 2022-03-02 code/benders.py
Tightened N in Benders cut and added toy_example_solve.py

@@ -244,14 +244,12 @@ class Benders:
         sub_path = [(e[0], e[1]) for e in T_E if self.sub_alpha[e[0], e[1]].X > 0.99]
 
         # some number larger than number of extra precedences in optimal master solution
-        N = self.instance.n * self.instance.n
         cut = self.master_model.addConstr(self.master_eta >= (self.sub_objval - self.LB) * gp.quicksum(
             1 / 3 * (self.master_y[e[0], e[1]] + self.master_x[e[0], master_M[e[0]]] +
                      self.master_x[e[1], master_M[e[1]]]) -
-            N * (3 - self.master_y[e[0], e[1]] - self.master_x[e[0], master_M[e[0]]] -
-                 self.master_x[e[1], master_M[e[1]]]) for e in sub_path)
-                                          - (self.sub_objval - self.LB) * (len(sub_path) - 1) + self.LB,
-                                          name="cut_{}".format(t))
+            (3 - self.master_y[e[0], e[1]] - self.master_x[e[0], master_M[e[0]]] - self.master_x[e[1], master_M[e[1]]])
+            for e in sub_path) - (self.sub_objval - self.LB) * (
+                                                  len(sub_path) - 1) + self.LB, name="cut_{}".format(t))
         # add cut to list of cuts
         self.cuts.append(cut)
 

### 2022-03-02 code/toy_example_solve.py

@@ -1,0 +1,51 @@
+import mrcpsp
+from compact_reformulation import CompactRefomulation
+from benders import Benders
+
+
+def solve_toy_example(nominal_data_file, Gamma, max_durational_deviations=None):
+    """
+    Solves toy example given in nominal_data_file for given Gamma. Max durational deviations can be explicitly provided
+    for each activity in the input, but if these are not provided, the max deviations are set to 0.5 * nominal values.
+    """
+    # load instance
+    instance = mrcpsp.load_nominal_mrcpsp(nominal_data_file)
+    if max_durational_deviations:
+        instance.set_dbar_explicitly(max_durational_deviations)
+    else:
+        instance.set_dbar_uncertainty_level(0.5)
+
+    print(f"\nSolving toy example given in {nominal_data_file}...")
+    print("\nJob data [r, d, d_bar]:")
+    print("-------------------------")
+    for i in instance.V:
+        print("job {}:".format(i),
+              [[instance.jobs[i].r[m][0], instance.jobs[i].d[m], instance.jobs[i].d_bar[m]] for m in
+               instance.jobs[i].M])
+
+    print("\nCompact reformulation")
+    print("---------------------")
+    compact_sol = CompactRefomulation(instance, Gamma, time_limit=20).solve(print_log=False)
+    print("objval:", compact_sol['objval'])
+    print("runtime:", compact_sol['runtime'])
+    print("modes:", compact_sol['modes'])
+    print("network:", compact_sol['network'])
+    print("resource flows:", compact_sol['flows'])
+
+    print("\nBenders'")
+    print("--------")
+    benders_sol = Benders(instance, Gamma, time_limit=20).solve(print_log=False)
+    print("objval:", benders_sol['objval'])
+    print("runtime:", benders_sol['runtime'])
+    print("modes:", benders_sol['modes'])
+    print("network:", compact_sol['network'])
+    print("resource flows:", benders_sol['flows'])
+
+
+if __name__ == "__main__":
+    balouka_toy_example = '/home/boldm1/OneDrive/robust-mrcpsp/instances/mrcpsp_toy_example_data.txt'
+    balouka_max_durational_deviations = {0: [0], 1: [0, 3], 2: [8, 2], 3: [1], 4: [3], 5: [0], 6: [0]}
+
+    paper_toy_example = '/home/boldm1/OneDrive/robust-mrcpsp/instances/paper_toy_example_data.txt'
+
+    solve_toy_example(balouka_toy_example, 3)

### 2022-03-02 code/balouka_toy_example.py
Removing balouka_toy_example.py

@@ -1,43 +1,0 @@
-
-import mrcpsp
-from compact_reformulation import CompactRefomulation
-from benders import Benders
-
-"""
-Testing code by solving toy example from Balouka and Cohen (2021).
-"""
-
-# read nominal data
-file_path = '/home/boldm1/OneDrive/robust-mrcpsp/instances/mrcpsp_toy_example_data.txt'
-instance = mrcpsp.load_nominal_mrcpsp(file_path)
-# set d_bar values explicitly
-instance.set_dbar_explicitly({0: [0], 1: [0, 3], 2: [8, 2], 3: [1], 4: [3], 5: [0], 6: [0]})
-
-# solve parameters
-Gamma = 3
-time_limit = 20
-
-print(f"\nSolving toy example given in {file_path}...")
-print("\nJob data [r, d, d_bar]:")
-print("-------------------------")
-for i in instance.V:
-    print("job {}:".format(i),
-          [[instance.jobs[i].r[m][0], instance.jobs[i].d[m], instance.jobs[i].d_bar[m]] for m in instance.jobs[i].M])
-
-print("\nCompact reformulation")
-print("---------------------")
-compact_sol = CompactRefomulation(instance, Gamma, time_limit).solve(print_log=False)
-print("objval:", compact_sol['objval'])
-print("runtime:", compact_sol['runtime'])
-print("modes:", compact_sol['modes'])
-print("network:", compact_sol['network'])
-print("resource flows:", compact_sol['flows'])
-
-print("\nBenders'")
-print("--------")
-benders_sol = Benders(instance, Gamma, time_limit).solve(print_log=False)
-print("objval:", benders_sol['objval'])
-print("runtime:", benders_sol['runtime'])
-print("modes:", benders_sol['modes'])
-print("network:", compact_sol['network'])
-print("resource flows:", benders_sol['flows'])

### 2022-03-02 code/benders.py
Constructing toy example.

@@ -94,10 +94,12 @@ class Benders:
                     master_objval = self.master_model.ObjVal
                     if master_objval > self.LB:
                         self.LB = self.master_model.ObjVal
+                    if print_log is True:
+                        self.print_master_sol()
                     # solve sub-problem and update UB
                     self.get_subproblem(t, num_threads, print_log)
                     if print_log is True:
-                        print("Solving sub-problem...")
+                        print("\nSolving sub-problem...")
                     self.sub_model.optimize()
                     self.sub_objval = self.sub_model.ObjVal
                     # if sub-problem provides tighter UB, update and record solution
@@ -129,6 +131,22 @@ class Benders:
                         self.add_cut(t)
                         iteration_times.append(time.perf_counter() - iteration_start)
                         t += 1
+
+    def print_master_sol(self):
+        """
+        Prints mode selection, transitive network and resource flows from master solution to the terminal.
+        """
+        modes = {i: int(sum(m * self.master_x[i, m].X for m in self.instance.jobs[i].M)) for i in
+                 self.instance.V}
+        T_E = [(i, j) for i in self.instance.V for j in self.instance.V if i != j if
+               self.master_y[i, j].X > 0.99]
+        flows = {(i, j): [self.master_f[i, j, k].X for k in self.instance.K_renew] for i in
+                 self.instance.V for j in self.instance.V if i != self.instance.n + 1 if j != 0 if
+                 i != j}
+        print("master objval:", self.master_model.ObjVal)
+        print("modes:", modes)
+        print("network:", T_E)
+        print("resource flows:", flows)
 
     def create_master_problem(self, num_threads=4, print_log=False):
         """

### 2022-03-02 code/mrcpsp.py

@@ -194,15 +194,26 @@ class Instance:
             specified for each possible mode.
         :type d_bar: dict
         """
+        # check each job has an entry in d_bar
+        no_value = []
+        for j in self.V:
+            try:
+                v = d_bar[j]
+            except KeyError:
+                no_value.append(j)
+        if len(no_value) > 0:
+            raise Exception(f"d_bar does not have an entry for job {str(no_value)[1:-1]}. Please correct this and try "
+                            "again.")
+
         # check each job in d_bar has values for each of its possible modes
         to_fix = []
         for j in self.V:
             if len(d_bar[j]) != len(self.jobs[j].M):
                 to_fix.append(j)
         # throw error if to_fix is non-empty
-        if to_fix:
-            raise Exception("The number of values specified in d_bar for jobs {} does not match the number of modes for"
-                            " those jobs! Please correct this and try again.".format(str(to_fix)[1:-1]))
+        if len(to_fix) > 0:
+            raise Exception(f"The number of values specified in d_bar for jobs {str(to_fix)[1:-1]} does not match the "
+                            "number of modes for those jobs! Please correct this and try again.")
 
         # set d_bar
         for j in d_bar:

### 2022-03-02 code/toy_example_solve.py

@@ -6,7 +6,7 @@ from benders import Benders
 def solve_toy_example(nominal_data_file, Gamma, max_durational_deviations=None):
     """
     Solves toy example given in nominal_data_file for given Gamma. Max durational deviations can be explicitly provided
-    for each activity in the input, but if these are not provided, the max deviations are set to 0.5 * nominal values.
+    for each activity in the input, but if not provided they are set to 0.5 * nominal values.
     """
     # load instance
     instance = mrcpsp.load_nominal_mrcpsp(nominal_data_file)
@@ -34,18 +34,26 @@ def solve_toy_example(nominal_data_file, Gamma, max_durational_deviations=None):
 
     print("\nBenders'")
     print("--------")
-    benders_sol = Benders(instance, Gamma, time_limit=20).solve(print_log=False)
+    benders_sol = Benders(instance, Gamma, time_limit=20).solve(print_log=True)
     print("objval:", benders_sol['objval'])
     print("runtime:", benders_sol['runtime'])
+    print("n_iterations:", benders_sol['n_iterations'])
     print("modes:", benders_sol['modes'])
     print("network:", compact_sol['network'])
     print("resource flows:", benders_sol['flows'])
 
 
 if __name__ == "__main__":
-    balouka_toy_example = '/home/boldm1/OneDrive/robust-mrcpsp/instances/mrcpsp_toy_example_data.txt'
+
+    # toy example from Balouka & Cohen (2021)
+    balouka_toy_example = '/home/boldm1/OneDrive/robust-mrcpsp/code/instances/mrcpsp_toy_example_data.txt'
     balouka_max_durational_deviations = {0: [0], 1: [0, 3], 2: [8, 2], 3: [1], 4: [3], 5: [0], 6: [0]}
 
-    paper_toy_example = '/home/boldm1/OneDrive/robust-mrcpsp/instances/paper_toy_example_data.txt'
+    # toy example from Bruni et al. (2017)
+    bruni_toy_example = '/home/boldm1/OneDrive/robust-mrcpsp/code/instances/rcpsp_toy_example_data.txt'
 
-    solve_toy_example(balouka_toy_example, 3)
+    # toy example for my paper
+    paper_toy_example = '/home/boldm1/OneDrive/robust-mrcpsp/code/instances/paper_toy_example_data.txt'
+    zero_durational_deviations = {0: [0], 1: [0], 2: [0, 0], 3: [0], 4: [0], 5: [0, 0], 6: [0]}
+
+    solve_toy_example(paper_toy_example, 2)

### 2022-03-03 code/benders.py
Developed toy example for paper.

@@ -80,6 +80,8 @@ class Benders:
                 self.master_model.setParam('TimeLimit', time_remaining)
                 # solve master problem
                 self.master_model.optimize()
+                if print_log is True:
+                    self.print_master_sol()
                 if self.master_model.status == 9:  # if time_limit reached whilst in master problem, terminate
                     # compute average iteration time, setting to +inf if terminated whilst in first iteration
                     if len(iteration_times) == 0:
@@ -94,14 +96,14 @@ class Benders:
                     master_objval = self.master_model.ObjVal
                     if master_objval > self.LB:
                         self.LB = self.master_model.ObjVal
-                    if print_log is True:
-                        self.print_master_sol()
                     # solve sub-problem and update UB
                     self.get_subproblem(t, num_threads, print_log)
                     if print_log is True:
                         print("\nSolving sub-problem...")
                     self.sub_model.optimize()
                     self.sub_objval = self.sub_model.ObjVal
+                    if print_log is True:
+                        self.print_subproblem_sol()
                     # if sub-problem provides tighter UB, update and record solution
                     if self.sub_objval < self.UB:
                         self.UB = self.sub_objval
@@ -147,6 +149,14 @@ class Benders:
         print("modes:", modes)
         print("network:", T_E)
         print("resource flows:", flows)
+
+    def print_subproblem_sol(self):
+        """
+        Prints longest path found in subproblem to the terminal.
+        """
+        T_E = [(i, j) for i in self.instance.V for j in self.instance.V if i != j if self.master_y[i, j].X > 0.99]
+        longest_path = [(e[0], e[1]) for e in T_E if self.sub_alpha[e[0], e[1]].X > 0.99]
+        print("longest path:", longest_path)
 
     def create_master_problem(self, num_threads=4, print_log=False):
         """

### 2022-03-03 code/toy_example_solve.py

@@ -51,6 +51,7 @@ if __name__ == "__main__":
 
     # toy example from Bruni et al. (2017)
     bruni_toy_example = '/home/boldm1/OneDrive/robust-mrcpsp/code/instances/rcpsp_toy_example_data.txt'
+    bruni_max_durational_deviations = {0: [0], 1: [2], 2: [1], 3: [3], 4: [2], 5: [1], 6: [1], 7: [1], 8: [0]}
 
     # toy example for my paper
     paper_toy_example = '/home/boldm1/OneDrive/robust-mrcpsp/code/instances/paper_toy_example_data.txt'

### 2022-03-03 code/experiment.py
Updating experiment.py and parallel_experiment.py to check for existing results file.

@@ -1,8 +1,10 @@
 import os
 import re
+import argparse
 from mrcpsp import load_nominal_mrcpsp
 from benders import Benders
 from compact_reformulation import CompactRefomulation
+from misc_functions import is_valid_directory
 
 
 def create_instances(instance_dir, uncertainty_level):
@@ -34,73 +36,96 @@ def create_instances(instance_dir, uncertainty_level):
     return instances
 
 
-def benders_experiment(instances, Gamma, time_limit):
+def run_experiment(instances, Gamma, solve_method, time_limit):
     """
-    Solves set uncertain MRCPSP instances using Bender' decomposition and writes results to an output file.
+    Solves a set of uncertain MRCPSP instances using specified solve method and writes results to an output file.
 
     :param instances: List of uncertain MRCPSP instances to solve.
     :type instances: list
     :param Gamma: Robustness parameter to use.
     :type Gamma: int
+    :param solve_method: Method to run experiment for. Either 'benders' or 'compact_reformulation'.
+    :type solve_method: str
     :param time_limit: Time limit in seconds to give to Benders' algorithm for each instance.
     :type time_limit: int
     """
-    # create results file
     current_dir = os.path.dirname(os.path.realpath(__file__))
-    results_file = os.path.join(current_dir, 'benders_results_{}.txt'.format(Gamma))
-    with open(results_file, 'w+') as f:
-        f.write("instance\tGamma\tobjval\tobjbound\tgap\tn_iterations\tavg_iteration\truntime\n")
 
-    # solve each instance using Benders' and write results
+    # Check if results already exist. Get solved instances if it does, or create new results file if not
+    results_file = os.path.join(current_dir, f'{solve_method}_results_{Gamma}.txt')
+    if os.path.exists(results_file):
+        with open(results_file, 'r') as f:
+            raw_lines = f.read().splitlines(keepends=True)
+        solved_instances = []
+        lines_to_keep = []
+        first_line = raw_lines[0]
+        for line in raw_lines[1:]:
+            splitline = line.split()
+            instance = splitline[0]
+            gap = float(splitline[4])
+            if solve_method == 'benders':
+                runtime = float(splitline[7])
+            elif solve_method == 'compact_reformulation':
+                runtime = float(splitline[5])
+            if (gap == 0) or (runtime > time_limit - 0.001):
+                solved_instances.append(instance)
+                lines_to_keep.append(line)
+        instances = [i for i in instances if i.name not in solved_instances]  # instances still to be solved
+        # re-write results file, removing partial solved instances
[…]

### 2022-03-03 code/misc_functions.py

@@ -1,0 +1,13 @@
+import os
+import argparse
+
+
+def is_valid_directory(arg):
+    """
+    Checks directory path given as input argument exists. Throws an error if it does not.
+    """
+    if os.path.isdir(arg):
+        return arg
+    else:
+        raise argparse.ArgumentTypeError("{} is not a valid directory.".format(arg))
+

### 2022-03-03 code/mrcpsp.py

@@ -190,8 +190,8 @@ class Instance:
         """
         Function to explicitly set d_bar for each mode of jobs in Instance.
 
-        :param d_bar: Dict with d_bar values for (a subset of) jobs in Instance. For each job in d_bar, values must be
-            specified for each possible mode.
+        :param d_bar: Dict with d_bar values for jobs in Instance. Values must be provided for each mode of each job in
+            instance.
         :type d_bar: dict
         """
         # check each job has an entry in d_bar

### 2022-03-03 code/parallel_experiment.py

@@ -7,6 +7,7 @@ from functools import partial
 from mrcpsp import load_nominal_mrcpsp
 from benders import Benders
 from compact_reformulation import CompactRefomulation
+from misc_functions import is_valid_directory
 
 
 def create_instances(instance_dir, uncertainty_level):
@@ -38,38 +39,49 @@ def create_instances(instance_dir, uncertainty_level):
     return instances
 
 
-def benders_solve_and_write(instance, Gamma, time_limit, num_threads):
+def solve_and_write(instance, Gamma, solve_method, time_limit, num_threads):
     """
-    Solves uncertain MRCPSP instance using Bender' decomposition and writes result to an (existing) output file.
+    Solves uncertain MRCPSP instance using specified solve method and writes result to an (existing) output file.
 
     :param instance: Uncertain MRCPSP instance to solve.
     :type instance: Instance
     :param Gamma: Robustness parameter to use.
     :type Gamma: int
+    :param solve_method: Method to use to solve instance. Either 'benders' or 'compact_reformulation'.
+    :type solve_method: str
     :param time_limit: Time limit in seconds to give to Benders' algorithm for each instance.
     :type time_limit: int
     :param num_threads: Number of threads to use when solving instance.
     :type num_threads: int
     """
     current_dir = os.path.dirname(os.path.realpath(__file__))
-    results_file = os.path.join(current_dir, 'benders_results_{}.txt'.format(Gamma))
+    results_file = os.path.join(current_dir, f'{solve_method}_results_{Gamma}.txt')
 
-    # solve instance using Benders' and write result
-    benders_sol = Benders(instance, Gamma, time_limit).solve(num_threads)
-    gap = abs(benders_sol['objbound'] - benders_sol['objval']) / abs(benders_sol['objval'])  # optimality gap
-    with FileLock(results_file + '.lock'):  # use FileLock to prevent two processes opening file at same time
-        with open(results_file, 'a') as f:
-            f.write("{}\t{}\t{:.3f}\t{:.3f}\t{:.3f}\t{}\t{:.3f}\t{:.3f}\n".format(instance.name, Gamma,
-                                                                                  benders_sol['objval'],
-                                                                                  benders_sol['objbound'], gap,
-                                                                                  benders_sol['n_iterations'],
-                                                                                  benders_sol['avg_iteration'],
-                                                                                  benders_sol['runtime']))
[…]

### 2022-03-03 code/plot_results.py

@@ -3,7 +3,7 @@ import numpy as np
 import matplotlib.pyplot as plt
 
 
-def read_results(results_dir, solve_methods, Gammas):
+def get_results(results_dir, solve_methods, Gammas):
     """
     Reads specified results files from directory and returns dictionary of results data.
 
@@ -36,10 +36,9 @@ def read_results(results_dir, solve_methods, Gammas):
     for method in solve_methods:
         results_data[method] = {}
         for g in Gammas:
-            file_name = f"{method}_results_{g}.txt"
-            file_path = os.path.join(results_dir, file_name)
+            file_name = os.path.join(results_dir, f"{method}_results_{g}.txt")
             file_data = {}
-            with open(file_path, 'r') as f:
+            with open(file_name, 'r') as f:
                 raw_lines = f.read().splitlines()
                 for line in raw_lines[1:]:
                     splitline = line.split()
@@ -52,7 +51,6 @@ def read_results(results_dir, solve_methods, Gammas):
                         file_data[instance] = {'objval': float(splitline[2]), 'objbound': float(splitline[3]),
                                                'gap': float(splitline[4]), 'runtime': float(splitline[5])}
             results_data[method][g] = file_data
-
     return results_data
 
 
@@ -186,7 +184,7 @@ def plot_performance_profile(results_data, solve_methods, Gammas, save_path):
 
 
 if __name__ == "__main__":
-    results = read_results('/home/boldm1/OneDrive/robust-mrcpsp/code/results/storm_j10_7200/',
+    results = get_results('/home/boldm1/OneDrive/robust-mrcpsp/code/results/storm_j10_7200/',
                            ['benders', 'compact_reformulation'], [3, 5, 7])
     plot_gaps(results, ['benders', 'compact_reformulation'], [3, 5, 7],
               '/home/boldm1/OneDrive/robust-mrcpsp/code/plots/')

### 2022-03-03 code/plot_results.py
Rerunning plots and creating latex table.

@@ -47,7 +47,7 @@ def get_results(results_dir, solve_methods, Gammas):
                         file_data[instance] = {'objval': float(splitline[2]), 'objbound': float(splitline[3]),
                                                'gap': float(splitline[4]), 'n_iterations': int(splitline[5]),
                                                'avg_iteration': float(splitline[6]), 'runtime': float(splitline[7])}
-                    elif method == 'compact_reformulation':
+                    elif 'compact_reformulation' in method:
                         file_data[instance] = {'objval': float(splitline[2]), 'objbound': float(splitline[3]),
                                                'gap': float(splitline[4]), 'runtime': float(splitline[5])}
             results_data[method][g] = file_data
@@ -104,9 +104,10 @@ def plot_gaps(results_data, solve_methods, Gammas, save_path):
     plt.rc('text', usetex=True)
     plt.figure()
     for method in solve_methods:
-        method_name = method.replace('_', ' ').capitalize()  # format method name for legend
+        method_name = method.replace('_', ' ').replace('trans', '').capitalize()  # format method name for legend
         plt.plot([100 * v for v in gap_values], [100 * v for v in less_than_gap[method]], label=method_name)
     plt.xlim([0, 100 * max_gap])
+    plt.ylim([0, 105])
     plt.xlabel(r'Gap (\%)')
     plt.ylabel(r'\% of instances')
     plt.legend(loc=4, frameon=False)
@@ -174,9 +175,10 @@ def plot_performance_profile(results_data, solve_methods, Gammas, save_path):
     plt.rc('text', usetex=True)
     plt.figure()
     for method in solve_methods:
-        method_name = method.replace('_', ' ').capitalize()  # format method name for legend
+        method_name = method.replace('_', ' ').replace('trans', '').capitalize()  # format method name for legend
         plt.plot(taus, perf_profile[method], label=method_name)
     plt.xlim([0, max(taus)])
+    plt.ylim([0, 1.05])
     plt.xlabel(r'$\tau$')
     plt.ylabel(r'$\mathrm{P}(\log(p_{im})\leq \tau)$')
     plt.legend(loc=4, frameon=False)
@@ -184,9 +186,9 @@ def plot_performance_profile(results_data, solve_methods, Gammas, save_path):
 
 
 if __name__ == "__main__":
[…]

### 2022-03-04 paper/paper.tex
Updates to paper.

@@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \documentclass[a4paper,abstracton]{scrartcl}
@@
 \graphicspath{{figures/}}  % path to directory containing all figures
+\usepackage{caption}  % allows subfigures
+\usepackage{subcaption}  % allows subfigures
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%  Macros
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% layout and formatting
+\allowdisplaybreaks  % allows page breaks within equations
+\usepackage{multirow}  % table formatting
+
+% macros
 \newcommand{\F}{{\mathcal{F}}}
@@
 
+* lit review *
+
 \section{Problem description}
@@
 Each mode for activity $i$, $m_i\in M_i$, has an associated renewable resource requirement of $r_{im_ik}$ for each $k\in K$, where $K$ is the set of renewable resource types involved in the project.
-Each renewable resource $k\in K$ has an availability of $R^_{k}$ at each time period in the project horizon.
+Each renewable resource $k\in K$ has an availability of $R_{k}$ at each time period in the project horizon.
 As well as resource constraints, the project is subject to a set of strict finish-to-start precedence constraints given by $E$, where $(i,j)\in E$ enforces that activity $i$ must finish before activity $j$ can begin.
 A project can be represented on a directed graph $G(V,E)$.
-/FIGURE shows an example project involving seven non-dummy activities and a single renewable resource.
+Figure \ref{fig:mrcpsp_network} shows an example uncertain MRCPSP instance involving five non-dummy activities and a single renewable resource.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[scale=1]{mrcpsp_network.pdf}
+	\caption{Example of an uncertain MRCPSP instance involving five non-dummy activities and a single renewable resource with an availability of 4.
+Activities 2 and 5 each have two available processing modes, whilst the other activities have only a single mode.}
+	\label{fig:mrcpsp_network}
+\end{figure}
 
 For a given choice of processing modes for each activity $\bm{m}$, we assume that the activity durations lie somewhere in a budgeted uncertainty set of the form
-$$\U_{\bm{m}}(\Gamma)=\Bigg\{\bm{d}\in\mathbb{R}_+^{|V|}:d_{im_i}=\bar{d}_{im_i}+\zeta_i\hat{d}_{im_i},\,0\leq \zeta_i \leq 1,\,\forall i\in V,\,\sum_{i\in V}\zeta_i \leq \Gamma\Bigg\}.$$
+$$\U_{\bm{m}}(\Gamma)=\Bigg\{\bm{d}\in\mathbb{R}_+^{|V|}:d_{im_i}=\bar{d}_{im_i}+\xi_i\hat{d}_{im_i},\,0\leq \xi_i \leq 1,\,\forall i\in V,\,\sum_{i\in V}\xi_i \leq \Gamma\Bigg\}.$$
 Introduced by \cite{bertsimas2004price}, the motivation of the budgeted uncertainty set is to control the pessimism of the solution by introducing a robustness parameter $\Gamma$ to limit the number of jobs that can simultaneously achieve their worst-case durations.
@@
 
-In this section we outline the Benders' decomposition approach to solving the robust MRCPSP first presented in \cite{balouka2021robust}.
+In this section we outline a Benders' decomposition approach to solving the robust MRCPSP first presented in \cite{balouka2021robust}.
+As mentioned previously, this approach is based on the approach used to solve the robust RCPSP in \cite{bruni2017adjustable}.
+
+The main idea behind this approach is to decompose the full problem into a two smaller problems: 1.
+a master problem that determines activity processing modes and resolves resource conflicts, and 2.
+a subproblem that takes the solution from the master problem and evaluates it by finding the worst-case makespan for the resulting network by solving a longest path problem.
+The solution to the master problem forms a lower bound to the optimal objective value of the original problem, whilst the solution to the subproblem provides an upper bound.
+These two problems are solved iteratively, with optimality cuts being added to the master problem each iteration based on the solution to the subproblem until the lower and upper bounds meet .
+
+\subsection{The master problem}
+
+The role of the master problem is to determine a choice of activity processing modes $\bm{m}$, as well as a sufficient selection $X_{\bm{m}}$ to resolve any resulting resource conflicts.
+The processing mode for each activity is determined by binary variables $x_{im}$.
+The $y_{ij}$ variables define a complete set of precedences between the project activities, either due to the original project precedence constraints, or due to the resolution of resource conflicts, i.e.
+$\{(i,j),\, i,j\in V:y_{ij}=1\}=T(E\cup X_{\bm{m}})$, where $T(E\cup X_{\bm{m}})$ is the transitive closure of the extended project precedence network.
+Resource flow variables $f_{ijk}$ track the amount of renewable resource $k$ that is transferred from activity $i$ to activity $j$ upon its completion.
+Finally, $s_i$ and $\phi_{ij}$ are additional variables used to define two relaxations of the subproblem to strengthen the master problem formulation.
+Using these variables, the master problem can be formulated as follows:
+\begin{align}
+	& \min \eta\label{eqn:master_first}\\
+	& y_{ij} = 1 \qquad \forall (i,j)\in E \label{eqn:master_y1}\\
+	& y_{ij} + y_{ji} \leq 1 \qquad \forall i,j\in V,\, i<j\label{eqn:master_y2}\\
+	& y_{ij} = y_{ip} + y_{p j} - 1 \qquad \forall i,j,p\in V,\,i\neq j\neq p\label{eqn:master_y3}\\
+	& f_{ijk} \leq \min(r_{im_ik}, r_{jm_jk})\cdot y_{ij} + (1 - x_{i,m_i})(\bar{r}_{ijk} - \min(r_{im_ik}, r_{jm_jk}))\nonumber\\
+	& & \hspace{-8cm} + (1 - x_{j,m_j})(\bar{r}_{ijk} - \min(r_{im_ik}, r_{jm_jk})) \nonumber \\
+	& & \hspace{-10cm}\forall i\in V\setminus\{n+1\},\, j \in V\setminus\{0\},\, i\neq j,\,m_i\in M_i,\,m_j\in M_j,\,k\in K\label{eqn:master_f1}\\
+	& \sum_{i\in V\setminus\{j,n+1\}}f_{ijk} = \sum_{m\in M_j}r_{jmk}x_{jm} \qquad \forall j \in V\setminus\{0\} ,\,k\in K\label{eqn:master_f2}\\
+	& \sum_{j\in V\setminus\{0,i\}}f_{ijk} = \sum_{m\in M_i}r_{imk}x_{im} \qquad \forall i \in V\setminus\{n+1\} ,\,k\in K\label{eqn:master_f3}\\
+	& \sum_{m\in M_i}x_{im} = 1\qquad \forall i \in V\label{eqn:master_x}\\
+	& \sum_{m\in M_i}r^{\nu}_{imk}x_{im} \leq R^{\nu}_k \qquad \forall i\in V,\,k\in K^{\nu}\label{eqn:master_nonrenewable}\\
+	& \eta \geq \lambda(x, y, M^{*\ell}) \qquad \forall \ell=0,\dots,t-1\label{eqn:master_cuts}\\
+	& s_j - s_i \geq \bar{d}_{im} - N(1-y_{ij}) - N(1-x_{im}) \qquad \forall i,j\in V,\, i\neq j,\,m\in M_i\label{eqn:master_relaxation1_1}\\
+	& \eta \geq s_{n+1}\label{eqn:master_relaxation1_2}\\
+	& \eta \geq \sum_{i \in V}\sum_{j\in V}\Big(\min_{m\in M_i}\bar{d}_{im}\Big)\phi_{ij}\label{eqn:master_relaxation2_1}\\
+	& \sum_{j\in V}\phi_{ji} = \sum_{j\in V}\phi_{ij} \qquad \forall i \in V\setminus\{0,n+1\}\\
+	& \sum_{j\in V}\phi_{0j} = 1\\
+	& \sum_{j\in V}\phi_{j,n+1} = 1\\
+	& \phi_{ij} \leq y_{ij} \qquad \forall i,j\in V,\,i\neq j\label{eqn:master_relaxation2_5}\\
+	& y_{ij} \in\{0,1\} \qquad \forall i,j\in V,\,i\neq j\\
+	& x_{im} \in \{0,1\} \qquad \forall i \in V,\, m\in M_i\\
+	& f_{ijk} \geq 0 \qquad \forall i\in V\setminus\{n+1\},\,j\in V\setminus\{0\},\,i\neq j,\,k\in K\\
+	& s_i \geq 0 \qquad \forall i\in V\\
+	& \phi_{ij} \geq 0 \qquad \forall i,j\in V \label{eqn:master_last}
+\end{align}
+where $\bar{r}_{ijk}=\max(\max_{m\in M_i}r_{imk},\,\max_{m\in M_j}r_{jmk})$ is the maximum usage of resource $k\in K$ by either activity $i$ or $j$, and as before, $N=\sum_{i\in V}\max_{m\in M_i}(\bar{d}_{im}+\hat{d}_{im})$ is an upper bound on the minimum makespan.
+
+The solution to he optimal objective value of the master problem (\ref{eqn:master_first})-(\ref{eqn:master_last}), denoted by $\eta^*$, forms a lower bound to the original robust MRCPSP problem.
+Note that for the first iteration, the solution to the master problem corresponds to an optimal solution to the MRCPSP with nominal activity durations.
+
+The original project precedence relations are captured by constraint (\ref{eqn:master_y1}).
+Constraints (\ref{eqn:master_y2}) and (\ref{eqn:master_y3}) ensure that the extended network is transitive and acyclic.
+Constraints (\ref{eqn:master_f1}) prevents the flow of resource from $i$ to $j$ from exceeding either the amount that becomes available from the completion of $i$, or the amount that is required by $j$, accounting for their selected modes.
+Constraints (\ref{eqn:master_f2}) and (\ref{eqn:master_f3}) ensure that each activity receives the amount of each renewable resource that it requires.
+Constraints (\ref{eqn:master_x}) allows only one processing mode to be selected for each activity.
+Non-renewable resource constraints are defined by (\ref{eqn:master_nonrenewable}).
+Constraints (\ref{eqn:master_cuts}) define the optimality cuts generated at each of the previous iterations.
+These cuts are explained in greater detail in Section \ref{section:optimality_cuts}.
+
+In line with \cite{bruni2017adjustable} and \cite{balouka2021robust}, two relaxations of the subproblem have been included in the master problem in order to strengthen it.
+The first relaxation is implemented through constraints (\ref{eqn:master_relaxation1_1}) and (\ref{eqn:master_relaxation1_2}).
+These simply enforce the extended precedence relationships using nominal activity durations for each processing mode.
+The second relaxation is captured by constraints (\ref{eqn:master_relaxation2_1})-(\ref{eqn:master_relaxation2_5}), which use flow variables $\phi_{ij}$ to form a longest-path problem through the extended project network using minimal activity durations.
+
+\subsection{The subproblem}
+
+The subproblem at iteration $t$ considers the worst-case makespan for the mode choices and sufficient selection from the master problem at iteration $t$, denoted by $\bm{m}^{*t}$ and $X_{\bm{m}^{*t}}$ respectively.
+This is done by solving an uncertain longest-path problem through the extended project network, where activity durations lie in the budgeted uncertainty set $\U_{\bm{m}^{*t}}(\Gamma)$.
+This problem can be formulated as
+\begin{align}
+V^{*t} = &\max \sum_{(i,j)\in T(E\cup X_{\bm{m}^{*t}})} \bar{d}_{im^{*t}_i}\alpha_{ij} + \hat{d}_{im^{*t}_i}w_{ij}\\
+	& \sum_{(i,n+1)\in T(E\cup X_{\bm{m}^{*t}})} \alpha_{i,n+1} = 1\\
+	& \sum_{(0,i)\in T(E\cup X_{\bm{m}^{*t}})} \alpha_{0,i} = 1\\
+	& \sum_{(i,j)\in T(E\cup X_{\bm{m}^{*t}})} \alpha_{ij} - \sum_{(j,i)\in T(E\cup X_{\bm{m}^{*t}})} \alpha_{ji} = 0 \qquad \forall i \in V\setminus\{0,n+1\}\\
+	& w_{ij} \leq \xi_i \qquad \forall (i,j)\in T(E\cup X_{\bm{m}^{*t}})\\
+	& w_{ij} \leq \alpha_{ij} \qquad \forall (i,j)\in T(E\cup X_{\bm{m}^{*t}})\\
+	& \sum_{i\in V}\xi_i \leq \Gamma\\
+	& \alpha_{ij}\in \{0,1\} \qquad \forall (i,j)\in T(E\cup X_{\bm{m}^{*t}})\\
+	& w_{ij}\geq 0 \qquad \forall (i,j)\in T(E\cup X_{\bm{m}^{*t}})\\
+	& 0\leq \xi_i \leq 1 \qquad \forall i\in V
+\end{align}
+
+Variables $\alpha_{ij}$ define the longest path through the network.
+This path, which we denote as $\pi^{*t}=\{(i,j),\,i,j\in V:\alpha_{ij}=1\}$ has length $V^{*t}$.
+Variables $\xi_i$ are the activity delay variables, whilst $w_{ij}$ are used to linearise the formulation.
+
+\subsection{Optimality cuts} \label{section:optimality_cuts}
+
+Following the solution to the subproblem, a valid cut is generated and added to the master problem for the next iteration.
+This cut simply forces an alternative solution in the master problem if it is to achieve a better objective value in the next iteration.
+
+Given the current best lower bound for the original problem $LB$, the mode selection from the master problem $\bm{m}^{*t}$, and the optimal objective value of the subproblem $V^{*t}$, the cut at iteration $t$ can be written as
+\begin{align}
+	& \eta \geq (V^{*t} - LB)\cdot\sum_{(i,j)\in \pi^{*t}}\bigg(1/3 (y_{ij} + x_{i,m_i^{*t}} + x_{j,m_j^{*t}}) - (3 - y_{ij} - x_{i,m_i^{*t}} - x_{j,m_j^{*t}})\bigg) & \nonumber \\
+	& \hspace{8.1cm} - (V^{*t} - LB)\cdot(|\pi^{*t}|-1) + LB,\label{eqn:benders_cut}
+\end{align}
+
+\cite{balouka2021robust} show that this constraint is a valid optimality cut for the problem, and that the number of these cuts that need to be added to the master problem before finding an optimal solution is finite.
+
+Before presenting a small example of the Benders' decomposition approach in the following subsection, we show an overview of the solution method in /ref.
+
+\subsection{Example}
 
 \section{Computational experiments and results}
+
+In this section we compare results from solving the compact reformulation with results from solving the Benders' decomposition approach.
+A complete set of the raw results used to generate the tables and plots presented here, in addition to the source code used to implement these experiments, can be found at \url{https://github.com/boldm1/robust-mrcpsp}.
+
+The instances used in this computational study have been created using the deterministic $j10$, and $j20$ MRCPSP instances from the PSPLIB (\cite{kolisch1997psplib}, \url{https://www.om-db.wi.tum.de/psplib/}).
+The $j10$ set contains a total of 536 instances each involving 10 activities, and the $j20$ set contains a total of 554 instances each involving 20 activities.
+We introduce uncertainty into these deterministic instances by setting the maximum deviation of the duration for each activity to be $\hat{d}_{im}=\lfloor0.7\times \bar{d}_{im}\rfloor$ for each mode $m\in M_i$.
+We then solve each instance for a range of robustness levels $\Gamma$.
+In particular we solve the $j10$ instances for $\Gamma\in\{0,3,5,7\}$ and the $j20$ instances for $\Gamma\in\{0,5,10,15\}$.
+
+Both the compact reformulation and Benders' decomposition approach were implemented in Python 3.9.2 and solved using Gurobi 9.0.1 running on a single core of a 2.30 GHz Intel Xeon CPU.
+A time limit of 2 hours per instance was imposed on each of the solution methods.
+
+Table \ref{table:benders_vs_compact} reports the percentage of instances solved to optimality (\textit{\%sol.}), the average optimality gap (in percent) over instances for which a feasible solution was found (\textit{gap}), and the average solution time (\textit{sol.
+time}) across the two instance sets and different choices of $\Gamma$, for both the Benders' decomposition approach and the compact reformulation.
+Note that if no optimal solution was found within the time limit by one of the solution methods, the solution time was recorded as the time limit value of 7200 seconds.
+Additionally, for the Benders' approach, the average number of iterations (\textit{it.}) and the average time per iteration (\textit{it.
+time}) are also reported.
+The average of these values across the different values for $\Gamma$ are also provided for each instance set.
+
+\begin{table}[h]
+\centering
+\small  % font size
+{\renewcommand{\arraystretch}{1.2}  % vertical padding
+\begin{tabular}{lrrrrrrrrrrrr}
+		     \hline \hline
+		     &      & & \multicolumn{5}{c}{Benders'}                                       & & \multicolumn{3}{c}{Compact reformulation} \\
+		     \cline{4-8} \cline{10-12}
+		     & $\Gamma$ & & \textit{\%sol.} & \textit{gap} & \textit{its.} & \textit{it. time} & \textit{sol. time} & & \textit{\%sol.}   & \textit{gap}   & \textit{sol. time}   \\
+		     \hline
+	\multirow{4}{*}{$j10$} & 0    & & 99.6     & 0.00     & 1            & 73.6          & 100.4        & & 100.00     & 0.00       & 1.8             \\
+			       & 3    & & 75.0     & 5.23     & 48           & 64.1          & 2065.6       & & 100.00     & 0.00       & 5.0             \\
+			       & 5    & & 72.8     & 6.53     & 53           & 67.9          & 2234.5       & & 100.00     & 0.00       & 4.6             \\
+			       & 7    & & 71.8     & 6.75     & 54           & 68.1          & 2278.4       & & 100.00     & 0.00       & 6.5             \\
+		     \cline{4-8} \cline{10-12}
+			       &      & & 79.8	   & 4.63     & 39.4	     & 68.4 	     & 1669.7	    & & 100.00     & 0.00	& 4.47 		  \\
+		     \hline
+	\multirow{4}{*}{$j20$} & 0    & & 80.8     & 0.00     & 1            & 123.5         & 1484.7       & & 89.4       & 1.92       & 941.9           \\
+			       & 5    & & 39.6     & 10.91    & 173          & 124.8         & 5181.7       & & 88.4       & 1.91       & 1061.3          \\
+			       & 10   & & 36.8     & 12.31    & 178          & 126.6         & 5406.2       & & 87.7       & 2.63       & 1138.4          \\
+			       & 15   & & 36.8     & 12.48    & 177          & 135.2         & 5415.8       & & 86.0       & 2.96       & 1279.8          \\
+		     \cline{4-8} \cline{10-12}
+			       &      & & 48.5	   & 8.93     & 132.5	     & 127.5	     & 4372.1       & & 87.9       & 2.36       & 1105.4          \\
+		     \hline \hline
+\end{tabular}
+}
+\caption{Comparison of the Benders' decomposition approach and the compact reformulation across the $j10$ and $j20$ instance sets and the different values of $\Gamma$.}
+\label{table:benders_vs_compact}
+\end{table}
+
+\begin{figure}[htb]
+	\makebox[\textwidth][c]{
+	\begin{subfigure}[t]{.65\textwidth}
+		\centering
+		\includegraphics[width=1.05\textwidth]{j10_performance_profile.pdf}
+		\caption{Performance profile of relative solution times.}\label{fig:j10_performance_profile}
+	\end{subfigure}%
+	\begin{subfigure}[t]{.65\textwidth}
+		\centering
+		\includegraphics[width=1.05\textwidth]{j10_gaps.pdf}
+		\caption{Percentage of instances solved to within given gap of optimality within the two-hour time limit.}\label{fig:j10_gaps}
+	\end{subfigure}
+	}
+	\caption{Comparison of Benders' approach and compact reformulation over instances in the j10 set.}
+	\label{fig:j10_plots}
+\end{figure}
+
+\begin{figure}[htb]
+	\makebox[\textwidth][c]{
+	\begin{subfigure}[t]{.65\textwidth}
+		\centering
+		\includegraphics[width=1.05\textwidth]{j20_performance_profile.pdf}
+		\caption{Performance profile of relative solution times.}\label{fig:j20_performance_profile}
+	\end{subfigure}%
+	\begin{subfigure}[t]{.65\textwidth}
+		\centering
+		\includegraphics[width=1.05\textwidth]{j20_gaps.pdf}
+		\caption{Percentage of instances solved to within given gap of optimality within the two-hour time limit.}\label{fig:j20_gaps}
+	\end{subfigure}
+	}
+	\caption{Comparison of Benders' approach and compact reformulation over instances in the j20 set.}
+	\label{fig:j20_plots}
+\end{figure}
 

### 2022-03-04 code/benders.py
Adding alternative 'compact' master problem to benders

@@ -36,6 +36,8 @@ class Benders:
         self.master_f = None
         self.master_s = None
         self.master_phi = None
+        self.master_S = None
+
         # Sub-problem model, objective value and parameters. Updated each time sub-problem is generated and solved.
         self.sub_model = None
         self.sub_objval = None
@@ -43,7 +45,7 @@ class Benders:
         self.sub_w = None
         self.sub_xi = None
 
-    def solve(self, num_threads=4, print_log=False):
+    def solve(self, num_threads=4, print_log=False, master_type='compact_base'):
         """
         Runs Benders' decomposition algorithm to solve robust MRCPSP instance. Terminates upon finding an optimal
         solution or reaching the specified time limit.
@@ -52,6 +54,8 @@ class Benders:
         :type num_threads: int
         :param print_log: Indicates whether or not to print Gurobi solve log to terminal. Defaults to False.
         :type print_log: bool
+        :param master_type: Master problem formulation to use. Either 'balouka' or 'compact'. Defaults to 'compact'.
+        :type str
         :return: Dictionary containing solution information.
         :rtype: dict
         """
@@ -60,7 +64,7 @@ class Benders:
         iteration_times = []  # List to store time of each iteration. Average iteration time is reported in solution.
         best_sol = {'modes': {}, 'network': [], 'flows': {}}  # dict to store x, y, f variable values of best solution
         start_time = time.perf_counter()
-        self.create_master_problem(num_threads, print_log)  # create basic master problem with no optimality cuts
+        self.create_master_problem(num_threads, print_log, master_type)
         while self.LB < self.UB:
             iteration_start = time.perf_counter()  # start timing iteration
             # set master problem time limit to remaining algorithm time limit to prevent algorithm getting stuck in the
@@ -158,11 +162,92 @@ class Benders:
         longest_path = [(e[0], e[1]) for e in T_E if self.sub_alpha[e[0], e[1]].X > 0.99]
         print("longest path:", longest_path)
 
-    def create_master_problem(self, num_threads=4, print_log=False):
+    def create_compact_master_problem(self, num_threads=4, print_log=False):
         """
         Creates Gurobi model to represent basic master problem without optimality cuts. The master problem selects job
[…]

### 2022-03-04 code/benders.py
Minor correction to create_compact_master_problem in benders

@@ -45,7 +45,7 @@ class Benders:
         self.sub_w = None
         self.sub_xi = None
 
-    def solve(self, num_threads=4, print_log=False, master_type='compact_base'):
+    def solve(self, num_threads=4, print_log=False, master_type='compact'):
         """
         Runs Benders' decomposition algorithm to solve robust MRCPSP instance. Terminates upon finding an optimal
         solution or reaching the specified time limit.
@@ -59,12 +59,20 @@ class Benders:
         :return: Dictionary containing solution information.
         :rtype: dict
         """
+        # check master type is recognised
+        if master_type not in ['balouka', 'compact']:
+            raise Exception("Specified master problem formulation type is not recognised. Must be either 'balouka' or "
+                            "'compact'. Please correct this and try again!")
+
         # Benders' decomposition algorithm
         t = 1  # iteration number
         iteration_times = []  # List to store time of each iteration. Average iteration time is reported in solution.
         best_sol = {'modes': {}, 'network': [], 'flows': {}}  # dict to store x, y, f variable values of best solution
         start_time = time.perf_counter()
-        self.create_master_problem(num_threads, print_log, master_type)
+        if master_type == 'compact':
+            self.create_compact_master_problem(num_threads, print_log)
+        elif master_type == 'balouka':
+            self.create_balouka_master_problem(num_threads, print_log)
         while self.LB < self.UB:
             iteration_start = time.perf_counter()  # start timing iteration
             # set master problem time limit to remaining algorithm time limit to prevent algorithm getting stuck in the
@@ -205,7 +213,7 @@ class Benders:
         model.setObjective(eta, GRB.MINIMIZE)
 
         # Precedence constraints
-        model.addConstr(S[0, 0] == 0)
+        model.addConstr(S[0] == 0)
         BigM = sum(max(d[i][m] + d_bar[i][m] for m in M[i]) for i in V)  # largest possible makespan
         model.addConstrs(
             S[j] - S[i] >= d[i][m] * x[i, m] - BigM * (1 - y[i, j]) for i in V for j in V for m in M[i])

### 2022-03-07 code/plot_results.py
Updating plots.

@@ -43,7 +43,7 @@ def get_results(results_dir, solve_methods, Gammas):
                 for line in raw_lines[1:]:
                     splitline = line.split()
                     instance = splitline[0]
-                    if method == 'benders':
+                    if 'benders' in method:
                         file_data[instance] = {'objval': float(splitline[2]), 'objbound': float(splitline[3]),
                                                'gap': float(splitline[4]), 'n_iterations': int(splitline[5]),
                                                'avg_iteration': float(splitline[6]), 'runtime': float(splitline[7])}
@@ -54,7 +54,7 @@ def get_results(results_dir, solve_methods, Gammas):
     return results_data
 
 
-def plot_gaps(results_data, solve_methods, Gammas, save_path):
+def plot_gaps(results_data, solve_methods, Gammas, save_path, legend=True):
     """
     Plots optimality gaps (x-axis) against the proportion of instances solved to within that gap (y-axis) for specified
     solve methods.
@@ -68,6 +68,8 @@ def plot_gaps(results_data, solve_methods, Gammas, save_path):
     :type Gammas: list
     :param save_path: Absolute filepath of directory in which to save plot.
     :type save_path: str
+    :param legend: Indicates whether or not to add legend to plot. Defaults to True.
+    :type legend: bool
     """
     # Check results_data has data for all combinations of solve method and Gamma
     missing_results = []
@@ -104,17 +106,18 @@ def plot_gaps(results_data, solve_methods, Gammas, save_path):
     plt.rc('text', usetex=True)
     plt.figure()
     for method in solve_methods:
-        method_name = method.replace('_', ' ').replace('trans', '').capitalize()  # format method name for legend
-        plt.plot([100 * v for v in gap_values], [100 * v for v in less_than_gap[method]], label=method_name)
+        plt.plot([100 * v for v in gap_values], [100 * v for v in less_than_gap[method]],
+                 color=get_solve_method_line_colour(method), label=get_solve_method_legend_name(method))
     plt.xlim([0, 100 * max_gap])
     plt.ylim([0, 105])
     plt.xlabel(r'Gap (\%)')
     plt.ylabel(r'\% of instances')
-    plt.legend(loc=4, frameon=False)
+    if legend is True:
+        plt.legend(loc=4, frameon=False)
     plt.savefig(os.path.join(save_path, 'gaps.pdf'))
 
 
[…]

### 2022-03-07 paper/paper.tex
Adding updated plots and table to paper.

@@
 \usepackage{bm}
+
+% algorithm presentation
+\usepackage{algpseudocode}
+\usepackage{algorithm}
+\algnewcommand{\Initialise}[1]{%
+	\State \textbf{Initialise:}\space\parbox[t]{.8\linewidth}{\raggedright #1}
+}
 
@@
 \usepackage{multirow}  % table formatting
+\usepackage{verbatim}  % allows blocks to be commented out
 
@@
 The multi-mode resource-constrained project scheduling problem (MRCPSP) is a generalisation of the widely-studied resource-constrained project scheduling problem (RCPSP) to include multiple processing modes for each activity.
-The purpose of introducing activity processing modes is to model situations in which there is more than one way of executing project activities, with each option having its own duration and resource requirements.
+The inclusion of these modes is to model situations in which there is more than one way of executing project activities, with each option having its own duration and resource requirements.
 The MRCPSP consists selecting the processing modes and start times for a given set of activities, subject to a set of precedence constraints and limited resource availability, with the objective of minimising the overall project duration, known as the makespan.
 
-The work in the paper extends the compact reformulation for the RCPSP developed by \cite{bold2021compact} for application to the MRCPSP.
-A computational study is presented comparing this approach against the Benders' decomposition approach used in \cite{balouka2021robust} for solving MRCPSP instances with uncertain activity durations.
-Note that the method used in \cite{balouka2021robust} is an extension of the approach used by \cite{bruni2017adjustable} to solve the RCPSP with uncertain activity durations.
+In this paper we consider the MRCPSP under uncertain activity durations and model it using a two-stage adjustable robust optimisation framework.
+In this setting, a first-stage problem is solved to determine activity mode selections and make activity sequencing decisions to resolve resource conflicts.
+Following this, the actual activity durations are realised and a complete schedule is computed.
+The aim of the two-stage robust MRCPSP is to find a feasible first-stage solution (i.e.
+mode selection and sequencing decisions) in order to minimise the realised worse-case makespan, as computed in the second-stage.
+We refer to this problem as the robust MRCPSP.
 
-* lit review *
+In recent years this two-stage robust optimisation approach has been applied to the RCPSP.
+First to use this approach were \cite{artigues2013robust}, who present an iterative scenario-relaxation algorithm for this problem with the objective of minimising the worst-case absolute regret.
+\cite{bruni2017adjustable} propose a Benders'-style decomposition approach for solving the robust RCPSP with the objective of minimising the worst-case project makespan.
+This work was extended in \cite{bruni2018computational}, which presents a computational study comparing an additional Benders' decomposition approach against a primal decomposition algorithm.
+Recently, \cite{bold2021compact} introduced a compact reformulation of the robust RCPSP and presented results which showed the superiority of that formulation over the iterative methods developed in those papers.
+
+The application of the this two-stage robust optimisation approach for the MRCPSP is a more recently development.
+To the best of our knowledge, the only existing paper to consider this problem is \cite{balouka2021robust}, in which the Benders' decomposition approach used by \cite{bruni2017adjustable} for the robust RCPSP has been extended for application to the MRCPSP.
+Analogous to that work, the aim of this paper is to apply the compact reformulation first developed for the RCPSP by \cite{bold2021compact} to the MRCPSP.
+Specifically, we present a computational comparison of our compact reformulation approach and a strengthened version of the Benders' decomposition approach.
 
@@
 Each activity $i\in V$ has a set of available processing modes given by $M_i=\{1,\dots,|M_i|\}$.
-We denote a particular choice of mode selections as $\bm{m}=(m_1,\dots,m_n)$, and the set of all possible mode selections by $\M\subseteq \mathbb{N}^n$.
-The nominal duration of activity $i$ when executed in mode $m_i\in M_i$ is given by $\hat{d}_{im_i}$, whilst its worst-case duration is given by $\bar{d}_{im_i}+\hat{d}_{im_i}$, where $\hat{d}_{im_i}$ is its maximum durational deviation.
+The nominal duration of activity $i$ when executed in mode $m_i\in M_i$ is given by $\bar{d}_{im_i}$, whilst its worst-case duration is given by $\bar{d}_{im_i}+\hat{d}_{im_i}$, where $\hat{d}_{im_i}$ is its maximum durational deviation.
 Each mode for activity $i$, $m_i\in M_i$, has an associated renewable resource requirement of $r_{im_ik}$ for each $k\in K$, where $K$ is the set of renewable resource types involved in the project.
 Each renewable resource $k\in K$ has an availability of $R_{k}$ at each time period in the project horizon.
-As well as resource constraints, the project is subject to a set of strict finish-to-start precedence constraints given by $E$, where $(i,j)\in E$ enforces that activity $i$ must finish before activity $j$ can begin.
-A project can be represented on a directed graph $G(V,E)$.
-Figure \ref{fig:mrcpsp_network} shows an example uncertain MRCPSP instance involving five non-dummy activities and a single renewable resource.
+As well as renewable resource requirements, mode $m_i$ of activity $i$ also has a non-renewable resource requirement of $r^{'}_{im_ik}$ for each non-renewable resource type $k\in K^{'}$, with each non-renewable resource having an overall availability of $R^{'}_k$ for the entire project horizon.
+Additionally, the project is subject to a set of strict finish-to-start precedence constraints given by $E$, where $(i,j)\in E$ enforces that activity $i$ must finish before activity $j$ can begin.
+These form a project network that can be represented using a directed graph $G(V,E)$.
+Figure \ref{fig:mrcpsp_network} shows an example instance involving five non-dummy activities and a single renewable resource.
 
@@
 	\includegraphics[scale=1]{mrcpsp_network.pdf}
-	\caption{Example of an uncertain MRCPSP instance involving five non-dummy activities and a single renewable resource with an availability of 4.
+	\caption{An example instance involving five non-dummy activities and a single renewable resource.
 Activities 2 and 5 each have two available processing modes, whilst the other activities have only a single mode.}
@@
 
-For a given choice of processing modes for each activity $\bm{m}$, we assume that the activity durations lie somewhere in a budgeted uncertainty set of the form
+For a given choice of processing modes for each activity $\bm{m}=(m_1,\dots,m_n)$, we assume that the activity durations lie somewhere in a budgeted uncertainty set of the form
 $$\U_{\bm{m}}(\Gamma)=\Bigg\{\bm{d}\in\mathbb{R}_+^{|V|}:d_{im_i}=\bar{d}_{im_i}+\xi_i\hat{d}_{im_i},\,0\leq \xi_i \leq 1,\,\forall i\in V,\,\sum_{i\in V}\xi_i \leq \Gamma\Bigg\}.$$
@@
 
-For a given selection of activity modes $\bm{m}$, a forbidden set is defined to be any subset of activities $F_{\bm{m}}\subseteq V$ that are not precedence-related, such that $\sum_{i\in F_{\bm{m}}}r_{im_ik}>R_k$ for at least one resource $k\in K$.
+For a given set of activity modes $\bm{m}$, a forbidden set is defined to be any subset of activities $F_{\bm{m}}\subseteq V$ that are not precedence-related, such that $\sum_{i\in F_{\bm{m}}}r_{im_ik}>R_k$ for at least one resource $k\in K$.
 That is, a forbidden set is a collection of activities that cannot be executed in parallel only because of resource limitations.
 
-Applying the main representation theorem of \cite{bartusch1988scheduling}, given a particular choice of activity modes $\bm{m}$, a solution to the MRCPSP can be defined by a set of additional precedences $X_{\bm{m}}\subseteq V^2\setminus E$ such that the extended precedence network $G(V,E\cup X_{\bm{m}})$ is acyclic and contains no forbidden sets.
+Applying the main representation theorem of \cite{bartusch1988scheduling}, for a particular choice of activity modes $\bm{m}$, a solution to the MRCPSP can be defined by a set of additional precedences $X_{\bm{m}}\subseteq V^2\setminus E$ such that the extended precedence network $G(V,E\cup X_{\bm{m}})$ is acyclic and contains no forbidden sets.
 Such an extension to the project network is referred to as a \textit{sufficient selection}.
@@
 \begin{align}
-	& \min_{\bm{m}\in \M,\,X_{\bm{m}}\in \X_{\bm{m}}} \max_{\bm{d}\in\U_{\bm{m}}(\Gamma)} \min_{\bm{S}} S_{n+1}\label{eqn:robust_mrcpsp_1}\\
+	& \min_{\bm{m}\in \M,\,X_{\bm{m}}\in \X_{\bm{m}}} \max_{\bm{d}\in\U_{\bm{m}}(\Gamma)} \min S_{n+1}\label{eqn:robust_mrcpsp_1}\\
 	& S_0 = 0\\
 	& S_j - S_i \geq d_{im_i} \qquad \forall (i,j)\in E\cup X_{\bm{m}}\\
-	& S_i \geq 0 \qquad \forall i\in V.\label{eqn:robust_mrcpsp_4}
+	& S_i \geq 0 \qquad \forall i\in V,\label{eqn:robust_mrcpsp_4}
 \end{align}
+where $\M\subseteq \mathbb{N}^n$ represents the set of all possible combinations of activity processing mode selections, and $\X_{\bm{m}}$ is the set of all possible sufficient selections for the choice of processing modes given by $\bm{m}$.
 This problem aims to selects activity modes and a corresponding sufficient selection in order to minimise the worst-case project makespan.
 
-\section{A compact reformulation}
+\section{A compact formulation}
 
-We propose solving this problem using an extended version of the compact reformulation from \cite{bold2021compact}.
-The development of this formulation for the RCPSP involved a reformulation of the adversarial sub-problem of determining the worst-case distribution of $\Gamma$ delays among the project activities.
-The reformulated sub-problem took the form of a longest-path problem through an augmented project network.
+We propose solving this problem using an extended version of the compact formulation developed by \cite{bold2021compact} for solving the robust RCPSP.
+This formulation combines the first and second-stage problems into a single compact formulation and was shown to significantly outperform the strongest iterative decompostion-based methods for the robust RCPSP.
 
-The derivation for the equivalent compact reformulation for the MRCPSP is more or less identical to the one presented in that paper for the RCPSP.
-We therefore omit its derivation here, and instead just present the resulting model.
+This formulation is constructed by recasting the adversarial subproblem of determining the worst-case activity durations for a given set of mode choices and sufficient selection as a longest path problem on an augmented project network.
+This augmented project network is formed of $\Gamma+1$ connected copies of the original project network, with each level being used to account for a delay to a single activity in the project.
+Having reformulated the adversarial subproblem as a longest-path problem, it can be dualised an inserted into a formulation for the full problem, to arrive at a complete compact formulation for the first-stage problem.
+Given that the derivation of this formulation for the MRCPSP is more or less identical to the derivation for the RCPSP, we omit it here and refer the reader to \cite{bold2021compact}.
 
+The resulting compact formulation for the two-stage adjustable robust MRCPSP can be written as
 \begin{align}
@@
 \text{s.t. } & S_{00}=0\label{eqn:compact2}\\
-	     & S_{j\gamma} - S_{i\gamma} \geq \bar{d}_{im}x_{im} - N(1-y_{ij}) & \hspace{-5cm}\forall (i,j)\in V^2,\,\forall m\in M_i,\,\gamma=0,\dots,\Gamma\label{eqn:compact_bigM1}\\
+	     & S_{j\gamma} - S_{i\gamma} \geq \bar{d}_{im}x_{im} - N(1-y_{ij}) \qquad \forall (i,j)\in V^2,\,\forall m\in M_i,\,\gamma=0,\dots,\Gamma\label{eqn:compact_bigM1}\\
 	     & S_{j,\gamma+1} - S_{i\gamma} \geq (\bar{d}_{im} + \hat{d}_{im})x_{im} - N(1-y_{ij})\nonumber \\
-	     &&\hspace{-5cm} \forall (i,j)\in V^2,\, m\in M_i,\,\gamma = 0,\dots,\Gamma-1\label{eqn:compact_bigM2}\\
-	     & y_{ij}=1 & \forall (i,j) \in E\cup\{(n+1,n+1)\}\label{eqn:compact5}\\
-	     & f_{ijk}\leq P_{ijk}y_{ij} & \hspace{-5cm}\forall (i,j)\in V^2,\,i\neq n+1,\,j\neq0,\, \forall k \in K\label{eqn:compact6}\\
-	     & \sum_{i\in V\setminus\{n+1\}}f_{ijk}=\sum_{m\in M_j}r_{jmk}x_{jm} & \forall j \in V\setminus\{0\} ,\, \forall k\in K\label{eqn:compact7}\\
-	     & \sum_{j\in V\setminus\{0\}}f_{ijk}=\sum_{m\in M_i}r_{imk}x_{im} & \forall i \in V\setminus\{n+1\},\,\forall k\in K\label{eqn:compact8}\\
-	     & \sum_{m\in M_i}x_{im} = 1 & \forall i\in V\label{eqn:compact9}\\
-	     & \sum_{m\in M_i}r^{\nu}_{imk}x_{im} \leq R^{\nu}_k & \forall i \in V,\,k\in K^{\nu}\\
-	     & S_{i\gamma}\geq 0 &\forall i\in V,\,\gamma \in 0,\dots, \Gamma\label{eqn:compact10}\\
-	     & f_{ijk}\geq 0 & \forall (i,j)\in V^2 ,\, \forall k \in K\label{eqn:compact11}\\
-	     & y_{ij}\in\{0,1\} & \forall (i,j)\in V^2\label{eqn:compact12}\\
-	     & x_{im}\in\{0,1\} & \forall i\in V,\,m\in M_i,
+	     & & \hspace{-8cm} \forall (i,j)\in V^2,\, m\in M_i,\,\gamma = 0,\dots,\Gamma-1\label{eqn:compact_bigM2}\\
+	     & y_{ij}=1 \qquad \forall (i,j) \in E\cup\{(n+1,n+1)\}\label{eqn:compact5}\\
+	     & y_{ij} + y_{ji} \leq 1 \qquad \forall i,j\in V,\, i<j\label{eqn:compact_trans1}\\
+	     & y_{ij} = y_{ip} + y_{p j} - 1 \qquad \forall i,j,p\in V,\,i\neq j\neq p\label{eqn:compact_trans2}\\
+	     & f_{ijk}\leq P_{ijk}y_{ij} \qquad \forall (i,j)\in V^2,\,i\neq n+1,\,j\neq0,\, \forall k \in K\label{eqn:compact6}\\
+	     & \sum_{i\in V\setminus\{n+1\}}f_{ijk}=\sum_{m\in M_j}r_{jmk}x_{jm} \qquad \forall j \in V\setminus\{0\} ,\, \forall k\in K\label{eqn:compact7}\\
+	     & \sum_{j\in V\setminus\{0\}}f_{ijk}=\sum_{m\in M_i}r_{imk}x_{im} \qquad \forall i \in V\setminus\{n+1\},\,\forall k\in K\label{eqn:compact8}\\
+	     & \sum_{m\in M_i}x_{im} = 1 \qquad \forall i\in V\label{eqn:compact9}\\
+	     & \sum_{m\in M_i}r^{'}_{imk}x_{im} \leq R^{'}_k \qquad \forall i \in V,\,k\in K^{'}\label{eqn:compact_nonrenewable}\\
+	     & S_{i\gamma}\geq 0 \qquad \forall i\in V,\,\gamma \in 0,\dots, \Gamma\label{eqn:compact10}\\
+	     & f_{ijk}\geq 0 \qquad \forall (i,j)\in V^2 ,\, \forall k \in K\label{eqn:compact11}\\
+	     & y_{ij}\in\{0,1\} \qquad \forall (i,j)\in V^2\label{eqn:compact12}\\
+	     & x_{im}\in\{0,1\} \qquad \forall i\in V,\,m\in M_i\label{eqn:compact13},
 \end{align}
@@
 
-This formulation can be solved directly by a standard mathematical optimisation solver such as Gurobi or CPLEX.
+Constraints (\ref{eqn:compact2})-(\ref{eqn:compact_bigM2}) are the dual constraints corresponding to the longest-path variables in the adversarial subproblem, and serve to ensure that the activity start times respect the project precedences as well as any delays to activity durations.
+Constraints (\ref{eqn:compact5}) capture the original project precedence relationships.
+Although not required for the correctness of the model, constraints (\ref{eqn:compact_trans1}) and (\ref{eqn:compact_trans2}) are additional transitivity constraints that have been included because they were shown in \cite{bold2021compact} to provide significant improvements to the computational performance of the model.
+Constraints (\ref{eqn:compact6})-(\ref{eqn:compact8}) are resource flow constraints which ensure that the transfer of resources between activities follows precedence constraints and that resources are conserved as they flow into and out of each activity in the network.
+Constraints (\ref{eqn:compact9}) allows only one processing mode to be selected for each activity.
+Finally, non-renewable resource constraints are enforced by (\ref{eqn:compact_nonrenewable}).
+
+With polynomially many constraints this formulation can be implemented straightforwardly using standard mathematical optimisation software such as Gurobi or CPLEX.
 
@@
 Using these variables, the master problem can be formulated as follows:
+\begin{align}
+\min \ & \eta\label{eqn:master_first}\\
+\text{s.t. } & \eta \geq S_{n+1}\\
+	     & S_{0}=0\label{eqn:master2}\\
+	     & S_{j} - S_{i} \geq \bar{d}_{im}x_{im} - N(1-y_{ij}) \qquad \forall (i,j)\in V^2,\,\forall m\in M_i\label{eqn:master3}\\
+	     & \textnormal{(\ref{eqn:compact5})-(\ref{eqn:compact_nonrenewable}),\,(\ref{eqn:compact11})-(\ref{eqn:compact13})}\\
+	     & S_i\geq 0 \qquad \forall i\in V\label{eqn:master_last}
+\end{align}
+
+\begin{comment}
 \begin{align}
@@
 where $\bar{r}_{ijk}=\max(\max_{m\in M_i}r_{imk},\,\max_{m\in M_j}r_{jmk})$ is the maximum usage of resource $k\in K$ by either activity $i$ or $j$, and as before, $N=\sum_{i\in V}\max_{m\in M_i}(\bar{d}_{im}+\hat{d}_{im})$ is an upper bound on the minimum makespan.
+\end{comment}
 
@@
 \begin{align}
-V^{*t} = &\max \sum_{(i,j)\in T(E\cup X_{\bm{m}^{*t}})} \bar{d}_{im^{*t}_i}\alpha_{ij} + \hat{d}_{im^{*t}_i}w_{ij}\\
+	V^{*t} = &\max \sum_{(i,j)\in T(E\cup X_{\bm{m}^{*t}})} \bar{d}_{im^{*t}_i}\alpha_{ij} + \hat{d}_{im^{*t}_i}w_{ij}\label{eqn:subproblem_first}\\
 	& \sum_{(i,n+1)\in T(E\cup X_{\bm{m}^{*t}})} \alpha_{i,n+1} = 1\\
@@
 	& w_{ij}\geq 0 \qquad \forall (i,j)\in T(E\cup X_{\bm{m}^{*t}})\\
-	& 0\leq \xi_i \leq 1 \qquad \forall i\in V
+	& 0\leq \xi_i \leq 1 \qquad \forall i\in V \label{eqn:subproblem_last}
 \end{align}
@@
 
-Before presenting a small example of the Benders' decomposition approach in the following subsection, we show an overview of the solution method in /ref.
+\begin{algorithm}[h] 
+\caption{Benders' decomposition algorithm}
+\begin{algorithmic}[1]
+\Initialise{Set $LB=-\infty$, $UB=+\infty$ and $t=1$.}
+\While{$UB > LB$}
+	\State \textbf{Solve master problem (\ref{eqn:master_first})-(\ref{eqn:master_last})} 
+	\State \hspace{\algorithmicindent} Get objective value $\eta^{*t}$, processing modes $\bm{m^{*t}}$ and precedences $\bm{y}_{\bm{m}^{*t}}$.
+	\State \hspace{\algorithmicindent} If $\eta^{*t}>LB$, update $LB\leftarrow \eta^{*t}$ 
+	\State \textbf{Solve subproblem (\ref{eqn:subproblem_first})-(\ref{eqn:subproblem_last})}
+	\State \hspace{\algorithmicindent} Get objective value $V^{*t}$ and longest path $\pi^{*t}$
+	\State \hspace{\algorithmicindent} If $V^{*t}<UB$, update $UB\leftarrow V^{*t}$
+	\State \textbf{Add cut (\ref{eqn:benders_cut}) to master problem}
+	\State Update $t\leftarrow t+1$
+\EndWhile
+\State \Return $UB$
+\end{algorithmic}
+\label{alg:benders}
+\end{algorithm}
 
 \subsection{Example}
+
+Here we solve the example shown in Figure \ref{fig:mrcpsp_network} with $R_1=4$ and $\Gamma=2$ to illustrate the implementation of the Benders' decomposition approach which is summarised in Algorithm \ref{alg:benders}.
+This problem is solved in 6 iterations, and the solution information from the master and subproblem at each of these iterations is shown in Table \ref{table:benders_example}.
+Note that in the final iteration, the algorithm can terminate after solving the master problem and updating the $UB$ since $LB=UB$, i.e.
+there is no need to solve the subproblem in the final iteration.
+
+\begin{table}[]
+\centering
+{\renewcommand{\arraystretch}{1.2}  % vertical padding
+\begin{tabular}{llllllll}
+	\hline \hline
+	$t$ & $LB$ & $UB$ & \eta^{*t} & $V^{*t}$ & \multicolumn{1}{c}{$\bm{m}^{*t}$}     & \multicolumn{1}{c}{$\pi^{*t}$}  \\
+	\hline
+1         & 11 & 16 & 11          & 16                                  & 1,2,1,1,2 & 0,1,2,5,6 \\
+2         & 12 & 15 & 12          & 15                                  & 1,1,1,1,1 & 0,1,3,2,4,6 \\
+3         & 12 & 15 & 12          & 15                                  & 1,1,1,1,1 & 0,1,2,3,4,6 \\
+4         & 13 & 15 & 13          & 18                                  & 1,2,1,1,1 & 0,1,2,5,6 \\
+5         & 13 & 15 & 13          & 16                                  & 1,1,1,2,1 & 0,1,3,4,2,5,6 \\
+6         & 15 & 15 & 15          & -                                   & 1,1,1,1,1 & - \\
+	\hline \hline
+\end{tabular}
+\caption{Iterations of the Benders' decomposition approach to solve example instance in Figure \ref{fig:mrcpsp_network}.}
+\label{table:benders_example}
+\end{table}
 
@@
 		     \hline \hline
-		     &      & & \multicolumn{5}{c}{Benders'}                                       & & \multicolumn{3}{c}{Compact reformulation} \\
+		     &      & & \multicolumn{5}{c}{Benders'}                                       & & \multicolumn{3}{c}{Compact formulation} \\
 		     \cline{4-8} \cline{10-12}
@@
 		     \hline
-	\multirow{4}{*}{$j10$} & 0    & & 99.6     & 0.00     & 1            & 73.6          & 100.4        & & 100.00     & 0.00       & 1.8             \\
-			       & 3    & & 75.0     & 5.23     & 48           & 64.1          & 2065.6       & & 100.00     & 0.00       & 5.0             \\
-			       & 5    & & 72.8     & 6.53     & 53           & 67.9          & 2234.5       & & 100.00     & 0.00       & 4.6             \\
-			       & 7    & & 71.8     & 6.75     & 54           & 68.1          & 2278.4       & & 100.00     & 0.00       & 6.5             \\
+	\multirow{4}{*}{$j10$} & 0    & & 100.0    & 0.00     & 1            & 0.96          & 1.1	    & & 100.0      & 0.00       & 1.8             \\
+			       & 3    & & 81.7     & 3.13     & 67           & 3.76          & 1551.4       & & 100.0      & 0.00       & 5.0             \\
+			       & 5    & & 79.7     & 4.20     & 76           & 3.89          & 1688.4       & & 100.0      & 0.00       & 4.6             \\
+			       & 7    & & 79.5     & 4.52     & 78           & 3.98          & 1701.1       & & 100.0      & 0.00       & 6.5             \\
 		     \cline{4-8} \cline{10-12}
-			       &      & & 79.8	   & 4.63     & 39.4	     & 68.4 	     & 1669.7	    & & 100.00     & 0.00	& 4.47 		  \\
+			       &      & & 85.2	   & 2.96     & 56	     & 3.15 	     & 1235.5	    & & 100.0      & 0.00	& 4.47 		  \\
 		     \hline
-	\multirow{4}{*}{$j20$} & 0    & & 80.8     & 0.00     & 1            & 123.5         & 1484.7       & & 89.4       & 1.92       & 941.9           \\
-			       & 5    & & 39.6     & 10.91    & 173          & 124.8         & 5181.7       & & 88.4       & 1.91       & 1061.3          \\
-			       & 10   & & 36.8     & 12.31    & 178          & 126.6         & 5406.2       & & 87.7       & 2.63       & 1138.4          \\
-			       & 15   & & 36.8     & 12.48    & 177          & 135.2         & 5415.8       & & 86.0       & 2.96       & 1279.8          \\
+	\multirow{4}{*}{$j20$} & 0    & & 89.7     & 0.00     & 1            & 176.62        & 901.3        & & 89.4       & 1.92       & 941.9           \\
+			       & 5    & & 51.3     & 10.04    & 228          & 135.66        & 4153.6       & & 88.4       & 1.91       & 1061.3          \\
+			       & 10   & & 50.7     & 10.81    & 228          & 140.37        & 4230.4       & & 87.7       & 2.63       & 1138.4          \\
+			       & 15   & & 49.8     & 11.05    & 222          & 148.72        & 4287.1       & & 86.0       & 2.96       & 1279.8          \\
 		     \cline{4-8} \cline{10-12}
-			       &      & & 48.5	   & 8.93     & 132.5	     & 127.5	     & 4372.1       & & 87.9       & 2.36       & 1105.4          \\
+			       &      & & 60.4	   & 7.98     & 170	     & 150.34	     & 3393.1       & & 87.9       & 2.36       & 1105.4          \\
 		     \hline \hline

### 2022-03-10 code/benders.py
Small changes to performance profile plots.

@@ -364,7 +364,6 @@ class Benders:
         # get longest path from sub-problem
         sub_path = [(e[0], e[1]) for e in T_E if self.sub_alpha[e[0], e[1]].X > 0.99]
 
-        # some number larger than number of extra precedences in optimal master solution
         cut = self.master_model.addConstr(self.master_eta >= (self.sub_objval - self.LB) * gp.quicksum(
             1 / 3 * (self.master_y[e[0], e[1]] + self.master_x[e[0], master_M[e[0]]] +
                      self.master_x[e[1], master_M[e[1]]]) -

### 2022-03-10 code/plot_results.py

@@ -180,12 +180,12 @@ def plot_performance_profile(results_data, solve_methods, Gammas, save_path, leg
     plt.rc('text', usetex=True)
     plt.figure()
     for method in solve_methods:
-        plt.plot(taus, perf_profile[method], color=get_solve_method_line_colour(method),
+        plt.plot(taus, [100 * p for p in perf_profile[method]], color=get_solve_method_line_colour(method),
                  label=get_solve_method_legend_name(method))
     plt.xlim([0, max(taus)])
-    plt.ylim([0, 1.05])
+    plt.ylim([0, 105])
     plt.xlabel(r'$\tau$')
-    plt.ylabel(r'$\mathrm{P}(\log(p_{im})\leq \tau)$')
+    plt.ylabel(r'\% of instances with $\log(r_{ia})\leq \tau$')
     if legend is True:
         plt.legend(loc=4, frameon=False)
     plt.savefig(os.path.join(save_path, 'performance_profile.pdf'))
@@ -221,9 +221,9 @@ def get_solve_method_legend_name(solve_method):
 
 
 if __name__ == "__main__":
-    results = get_results('/home/boldm1/OneDrive/robust-mrcpsp/code/results/storm_j10_7200/',
-                          ['benders', 'new_benders', 'compact_reformulation_trans'], [3, 5, 7])
-    plot_gaps(results, ['benders', 'new_benders', 'compact_reformulation_trans'], [3, 5, 7],
-              '/home/boldm1/OneDrive/robust-mrcpsp/code/plots/')
-    plot_performance_profile(results, ['benders', 'new_benders', 'compact_reformulation_trans'], [3, 5, 7],
-                             '/home/boldm1/OneDrive/robust-mrcpsp/code/plots/', legend=False)
+    results_folder = '/home/boldm1/OneDrive/robust-mrcpsp/code/results/storm_j20_7200/'
+    methods = ['benders', 'new_benders', 'compact_reformulation_trans']
+    Gammas = [5, 10, 15]
+    results = get_results(results_folder, methods, Gammas)
+    plot_gaps(results, methods, Gammas, '/home/boldm1/OneDrive/robust-mrcpsp/code/plots/')
+    plot_performance_profile(results, methods, Gammas, '/home/boldm1/OneDrive/robust-mrcpsp/code/plots/', legend=False)

### 2022-03-10 paper/paper.tex
Updates to paper.

@@
 \bibliographystyle{apalike}
+\usepackage{bibentry}  % inline referencing
+\renewcommand{\inlinebibentry}[1]{\cite{#1} \bibentry{#1}}
 \usepackage{hyperref}
@@
 \author[2]{Marc Goerigk}
-
 \affil[1]{STOR-i Centre for Doctoral Training, Lancaster University, Lancaster, UK}	
@@
 
+\nobibliography*  % tells bibentry to load bibliography database
+
 \begin{abstract}
+	This paper presents a mixed-integer linear programming formulation for the multi-mode resource-constrained project scheduling problem with uncertain activity durations.
+We consider a two-stage robust optimisation approach, where activity modes are selected and resource conflicts are resolved under the uncertain activity durations in a first-stage, which is followed by a second-stage where the schedule is evaluated once the activity durations become known.
+We aim to find solutions to minimise the worst-case project makespan, whilst assuming that activity durations lie in a budgeted uncertainty set.
+
+	The proposed mixed-integer linear programming formulation is an extension of the formulation proposed by \cite{bold2021compact} (\bibentry{bold2021compact}) for the problem without multiple activity modes.
+Computational experiments show that this easy-to-implement formulation can be solved many times faster than the current state-of-the-art for solving this problem, solving over 40\% more instances to optimality over the same set of benchmarking instances.
 \end{abstract}
@@
 
-We propose solving this problem using an extended version of the compact formulation developed by \cite{bold2021compact} for solving the robust RCPSP.
-This formulation combines the first and second-stage problems into a single compact formulation and was shown to significantly outperform the strongest iterative decompostion-based methods for the robust RCPSP.
+We propose solving the robust MRCPSP using an extended version of the mixed-integer programming formulation developed by \cite{bold2021compact} for solving the robust RCPSP.
+This formulation combines the first and second-stage problems into a single compact formulation and was shown to significantly outperform the strongest iterative decomposition-based methods for the robust RCPSP.
 
@@
 This augmented project network is formed of $\Gamma+1$ connected copies of the original project network, with each level being used to account for a delay to a single activity in the project.
-Having reformulated the adversarial subproblem as a longest-path problem, it can be dualised an inserted into a formulation for the full problem, to arrive at a complete compact formulation for the first-stage problem.
-Given that the derivation of this formulation for the MRCPSP is more or less identical to the derivation for the RCPSP, we omit it here and refer the reader to \cite{bold2021compact}.
+Having reformulated the adversarial subproblem as a longest-path problem, it can be dualised an inserted into a formulation for the first-stage problem, to arrive at a complete compact formulation for the full problem.
+Given that the derivation of this formulation is more or less identical to the derivation of the corresponding formulation for the robust RCPSP, we omit it here and instead refer the reader to \cite{bold2021compact}.
 
-The resulting compact formulation for the two-stage adjustable robust MRCPSP can be written as
+In the resulting compact formulation, $y_{ij}$ variables are used to define a complete set of transitive precedences between the project activities that are formed by the original project precedence constraints and the additional precedences introduced to resolve resource conflicts, i.e.
+$\{(i,j),\, i,j\in V:y_{ij}=1\}=T(E\cup X_{\bm{m}})$, where $T(\cdot)$ is used to denote the transitive closure of a set.
+Variables $x_{im}$ determine the processing mode for each activity, whilst resource flow variables $f_{ijk}$ track the amount of renewable resource $k$ that is transferred from activity $i$ to activity $j$ upon its completion.
+The $S_{i\gamma}$ variables are dual variables associated with the longest-path subproblem, which in this formulation these variables are used to track the start time of activities under the worst-case activity durations.
+Using these variables, the compact formulation for the robust MRCPSP can be written as
 \begin{align}
@@
 	     & S_{i\gamma}\geq 0 \qquad \forall i\in V,\,\gamma \in 0,\dots, \Gamma\label{eqn:compact10}\\
-	     & f_{ijk}\geq 0 \qquad \forall (i,j)\in V^2 ,\, \forall k \in K\label{eqn:compact11}\\
-	     & y_{ij}\in\{0,1\} \qquad \forall (i,j)\in V^2\label{eqn:compact12}\\
+	     & y_{ij}\in\{0,1\} \qquad \forall (i,j)\in V^2\label{eqn:compact11}\\
+	     & f_{ijk}\geq 0 \qquad \forall (i,j)\in V^2 ,\, \forall k \in K\label{eqn:compact12}\\
 	     & x_{im}\in\{0,1\} \qquad \forall i\in V,\,m\in M_i\label{eqn:compact13},
@@
 
-With polynomially many constraints this formulation can be implemented straightforwardly using standard mathematical optimisation software such as Gurobi or CPLEX.
+With polynomially many variables and constraints this formulation can be implemented straightforwardly using standard mathematical optimisation software such as Gurobi or CPLEX.
 
@@
 
-In this section we outline a Benders' decomposition approach to solving the robust MRCPSP first presented in \cite{balouka2021robust}.
-As mentioned previously, this approach is based on the approach used to solve the robust RCPSP in \cite{bruni2017adjustable}.
+In this section we outline an alternative approach to solving the robust MRCPSP based on Benders' decomposition.
+Such a solution method for the robust MRCPSP was first presented in \cite{balouka2021robust}, and as mentioned previously, this itself was based on the approach to solve the robust RCPSP used in \cite{bruni2017adjustable}.
 
@@
 The solution to the master problem forms a lower bound to the optimal objective value of the original problem, whilst the solution to the subproblem provides an upper bound.
-These two problems are solved iteratively, with optimality cuts being added to the master problem each iteration based on the solution to the subproblem until the lower and upper bounds meet .
+These two problems are solved iteratively, with optimality cuts being added to the master problem each iteration based on the solution to the subproblem until the lower and upper bounds meet.
+
+After initially implementing the Benders' algorithm as it is presented in \cite{balouka2021robust}, it was discovered that its performance could be strengthened by replacing the master problem formulation used in that implementation with a stripped-down version of the compact formulation (\ref{eqn:compact1})-(\ref{eqn:compact13}) in which the uncertainty is removed.
+In this section we therefore present our strengthened implementation of the Benders' approach for solving the robust MRCPSP.
+Note that a comparison of these two Benders' implementations is included in Section \ref{section:experiments}.
+For the details of the original implementation of the Benders' decomposition algorithm for the robust MRCPSP, see \cite{balouka2021robust}.
 
@@
 The role of the master problem is to determine a choice of activity processing modes $\bm{m}$, as well as a sufficient selection $X_{\bm{m}}$ to resolve any resulting resource conflicts.
-The processing mode for each activity is determined by binary variables $x_{im}$.
-The $y_{ij}$ variables define a complete set of precedences between the project activities, either due to the original project precedence constraints, or due to the resolution of resource conflicts, i.e.
-$\{(i,j),\, i,j\in V:y_{ij}=1\}=T(E\cup X_{\bm{m}})$, where $T(E\cup X_{\bm{m}})$ is the transitive closure of the extended project precedence network.
-Resource flow variables $f_{ijk}$ track the amount of renewable resource $k$ that is transferred from activity $i$ to activity $j$ upon its completion.
-Finally, $s_i$ and $\phi_{ij}$ are additional variables used to define two relaxations of the subproblem to strengthen the master problem formulation.
-Using these variables, the master problem can be formulated as follows:
+This problem is solved without the direct consideration of the uncertain activity durations, and instead, uncertainty is accounted for in the subproblem and communicated back to the master problem with the use of optimality cuts.
+Hence, to remove the uncertainty from the model, (\ref{eqn:compact1})-(\ref{eqn:compact13}) is modified by considering only a single level of the augmented project network.
+Using this formulation, the master problem can be written as
 \begin{align}
@@
 	     & S_{j} - S_{i} \geq \bar{d}_{im}x_{im} - N(1-y_{ij}) \qquad \forall (i,j)\in V^2,\,\forall m\in M_i\label{eqn:master3}\\
+	     & \eta \geq \lambda(x, y, M^{*\ell}) \qquad \forall \ell=1,\dots,t-1\label{eqn:master_cuts}\\
 	     & \textnormal{(\ref{eqn:compact5})-(\ref{eqn:compact_nonrenewable}),\,(\ref{eqn:compact11})-(\ref{eqn:compact13})}\\
-	     & S_i\geq 0 \qquad \forall i\in V\label{eqn:master_last}
+	     & S_i\geq 0 \qquad \forall i\in V,\label{eqn:master_last}
 \end{align}
+where (\ref{eqn:master_cuts}) are the optimality cuts generated from the solutions of the resulting subproblems at each of the previous iterations.
 
+The optimal objective value of the master problem (\ref{eqn:master_first})-(\ref{eqn:master_last}), denoted by $\eta^*$, forms a lower bound to the original robust MRCPSP problem.
+Note that for the first iteration, the solution to the master problem corresponds to an optimal solution to the MRCPSP with nominal activity durations.
+
+% Master problem formulation from Balouka & Cohen (2021)
 \begin{comment}
+The processing mode for each activity is determined by binary variables $x_{im}$.
+The $y_{ij}$ variables define a complete set of precedences between the project activities, either due to the original project precedence constraints, or due to the resolution of resource conflicts, i.e.
+$\{(i,j),\, i,j\in V:y_{ij}=0\}=T(E\cup X_{\bm{m}})$, where $T(E\cup X_{\bm{m}})$ is the transitive closure of the extended project precedence network.
+Resource flow variables $f_{ijk}$ track the amount of renewable resource $k$ that is transferred from activity $i$ to activity $j$ upon its completion.
+Finally, $s_i$ and $\phi_{ij}$ are additional variables used to define two relaxations of the subproblem to strengthen the master problem formulation.
+Using these variables, the master problem can be formulated as follows:
+
 \begin{align}
@@
 where $\bar{r}_{ijk}=\max(\max_{m\in M_i}r_{imk},\,\max_{m\in M_j}r_{jmk})$ is the maximum usage of resource $k\in K$ by either activity $i$ or $j$, and as before, $N=\sum_{i\in V}\max_{m\in M_i}(\bar{d}_{im}+\hat{d}_{im})$ is an upper bound on the minimum makespan.
-\end{comment}
-
-The solution to he optimal objective value of the master problem (\ref{eqn:master_first})-(\ref{eqn:master_last}), denoted by $\eta^*$, forms a lower bound to the original robust MRCPSP problem.
-Note that for the first iteration, the solution to the master problem corresponds to an optimal solution to the MRCPSP with nominal activity durations.
 
@@
 The second relaxation is captured by constraints (\ref{eqn:master_relaxation2_1})-(\ref{eqn:master_relaxation2_5}), which use flow variables $\phi_{ij}$ to form a longest-path problem through the extended project network using minimal activity durations.
+\end{comment}
 
@@
 The subproblem at iteration $t$ considers the worst-case makespan for the mode choices and sufficient selection from the master problem at iteration $t$, denoted by $\bm{m}^{*t}$ and $X_{\bm{m}^{*t}}$ respectively.
-This is done by solving an uncertain longest-path problem through the extended project network, where activity durations lie in the budgeted uncertainty set $\U_{\bm{m}^{*t}}(\Gamma)$.
-This problem can be formulated as
+This is done by solving an uncertain longest-path problem through the extended project network $T(E\cup X_{\bm{m}^{*t}})$, where activity durations lie in the budgeted uncertainty set $\U_{\bm{m}^{*t}}(\Gamma)$, which can be formulated as
 \begin{align}
@@
 	& w_{ij}\geq 0 \qquad \forall (i,j)\in T(E\cup X_{\bm{m}^{*t}})\\
-	& 0\leq \xi_i \leq 1 \qquad \forall i\in V \label{eqn:subproblem_last}
+	& 0\leq \xi_i \leq 1 \qquad \forall i\in V. \label{eqn:subproblem_last}
 \end{align}
@@
 Variables $\alpha_{ij}$ define the longest path through the network.
-This path, which we denote as $\pi^{*t}=\{(i,j),\,i,j\in V:\alpha_{ij}=1\}$ has length $V^{*t}$.
+This path, which is denote as $\pi^{*t}=\{(i,j),\,i,j\in V:\alpha_{ij}=1\}$ has length $V^{*t}$.
 Variables $\xi_i$ are the activity delay variables, whilst $w_{ij}$ are used to linearise the formulation.
+Note that this formulation of the subproblem is identical to the one implemented by \cite{balouka2021robust}.
 
@@
 
+\subsection{Example}
+
+Here, to demonstrate its implementation, we use the Benders' decomposition approach to solve the example shown in Figure \ref{fig:mrcpsp_network} with $R_1=4$ and $\Gamma=2$.
+An outline of the implementation of the Benders' approach is presented in Algorithm \ref{alg:benders}.
+
+The algorithm solves this instance in 6 iterations, and the solution information from the master and subproblem at each of these iterations is shown in Table \ref{table:benders_example}.
+Note that in the final iteration, the algorithm can terminate after solving the master problem and updating the $UB$ since $LB=UB$, i.e.
+there is no need to solve the subproblem in the final iteration.
+
 \begin{algorithm}[h] 
-\caption{Benders' decomposition algorithm}
+\caption{Benders' decomposition algorithm.}
 \begin{algorithmic}[1]
@@
 	\State \textbf{Solve master problem (\ref{eqn:master_first})-(\ref{eqn:master_last})} 
-	\State \hspace{\algorithmicindent} Get objective value $\eta^{*t}$, processing modes $\bm{m^{*t}}$ and precedences $\bm{y}_{\bm{m}^{*t}}$.
+	\State \hspace{\algorithmicindent} Get objective value $\eta^{*t}$, processing modes $\bm{m^{*t}}$ and precedences $\bm{y}_{\bm{m}^{*t}}$
 	\State \hspace{\algorithmicindent} If $\eta^{*t}>LB$, update $LB\leftarrow \eta^{*t}$ 
@@
 
-\subsection{Example}
-
-Here we solve the example shown in Figure \ref{fig:mrcpsp_network} with $R_1=4$ and $\Gamma=2$ to illustrate the implementation of the Benders' decomposition approach which is summarised in Algorithm \ref{alg:benders}.
-This problem is solved in 6 iterations, and the solution information from the master and subproblem at each of these iterations is shown in Table \ref{table:benders_example}.
-Note that in the final iteration, the algorithm can terminate after solving the master problem and updating the $UB$ since $LB=UB$, i.e.
-there is no need to solve the subproblem in the final iteration.
-
-\begin{table}[]
+\begin{table}[h]
 \centering
@@
 \end{tabular}
-\caption{Iterations of the Benders' decomposition approach to solve example instance in Figure \ref{fig:mrcpsp_network}.}
+\caption{Solution information at each of the six iterations of the Benders' decomposition approach required to solve example instance in Figure \ref{fig:mrcpsp_network}.}
 \label{table:benders_example}
@@
 
-\section{Computational experiments and results}
+\section{Computational experiments and results} \label{section:experiments}
 
-In this section we compare results from solving the compact reformulation with results from solving the Benders' decomposition approach.
+In this section we compare results from using the compact formulation and the Benders' decomposition approach to solve uncertain MRCPSP instances.
 A complete set of the raw results used to generate the tables and plots presented here, in addition to the source code used to implement these experiments, can be found at \url{https://github.com/boldm1/robust-mrcpsp}.
@@
 The $j10$ set contains a total of 536 instances each involving 10 activities, and the $j20$ set contains a total of 554 instances each involving 20 activities.
-We introduce uncertainty into these deterministic instances by setting the maximum deviation of the duration for each activity to be $\hat{d}_{im}=\lfloor0.7\times \bar{d}_{im}\rfloor$ for each mode $m\in M_i$.
-We then solve each instance for a range of robustness levels $\Gamma$.
+We introduce uncertainty into these deterministic instances by setting the maximum durational deviation for each activity to be $\hat{d}_{im}=\lfloor0.7\times \bar{d}_{im}\rfloor$ for each mode $m\in M_i$.
+These uncertain instances have then been solved using both methods for a range of robustness levels $\Gamma$.
 In particular we solve the $j10$ instances for $\Gamma\in\{0,3,5,7\}$ and the $j20$ instances for $\Gamma\in\{0,5,10,15\}$.
 
-Both the compact reformulation and Benders' decomposition approach were implemented in Python 3.9.2 and solved using Gurobi 9.0.1 running on a single core of a 2.30 GHz Intel Xeon CPU.
+Both the compact reformulation and Benders' decomposition approach have been implemented in Python 3.9.2 and solved using Gurobi 9.0.1 running on a single core of a 2.30 GHz Intel Xeon CPU.
 A time limit of 2 hours per instance was imposed on each of the solution methods.
 
-Table \ref{table:benders_vs_compact} reports the percentage of instances solved to optimality (\textit{\%sol.}), the average optimality gap (in percent) over instances for which a feasible solution was found (\textit{gap}), and the average solution time (\textit{sol.
-time}) across the two instance sets and different choices of $\Gamma$, for both the Benders' decomposition approach and the compact reformulation.
+We begin by observing the effect of the robustness parameter $\Gamma$ on the average optimal objective values shown in Tables \ref{table:j10_objvals} and \ref{table:j20_objvals}.
+The values in these tables have been computed only over the instances for which an optimal solution was found for all the values of $\Gamma$ (i.e.
+all 536 instances in the $j10$ set and 477 instances of the $j20$ set).
+As expected, as $\Gamma$ increases the solution cost increases in a concave manner.
+
+\begin{table}[h]
+	\centering
+	{\renewcommand{\arraystretch}{1.2}  % vertical padding
+	\begin{subtable}{.45\linewidth}
+	\begin{tabular}{lllll}
+		\hline\hline
+		$\Gamma$  & 0     & 3     & 5     & 7 \\
+		\hline
+		obj. & 16.84 & 25.34 & 26.35 & 26.46 \\               
+		\hline\hline
+	\end{tabular}
+	\caption{$j10$}
+	\label{table:j10_objvals}
+	\end{subtable}%
+	\qquad
+	\begin{subtable}{.45\linewidth}
+	\begin{tabular}{lllll}
+		\hline\hline
+		$\Gamma$  & 0     & 5     & 10     & 15 \\
+		\hline
+		obj. & 24.61 & 37.89 & 38.69 & 38.69 \\               
+		\hline\hline
+	\end{tabular}
+	\caption{$j20$}
+	\label{table:j20_objvals}
+	\end{subtable}
+	\caption{Average optimal objective values across instances in the $j10$ and $j20$ instance sets for different values of $\Gamma$.}
+	\label{table:objvals}
+\end{table}
+
+Table \ref{table:benders_vs_compact} reports the percentage of instances solved to optimality (\textit{\%sol.}), the average percentage optimality gap over instances for which a feasible solution was found (\textit{gap}), and the average solution time in seconds (\textit{sol.
+time}) for the different choices of $\Gamma$ across the two instance sets, for both the Benders' decomposition approach and the compact formulation.
 Note that if no optimal solution was found within the time limit by one of the solution methods, the solution time was recorded as the time limit value of 7200 seconds.
-Additionally, for the Benders' approach, the average number of iterations (\textit{it.}) and the average time per iteration (\textit{it.
+Additionally, for the Benders' approach, the average number of (completed) iterations (\textit{it.}) and the average time per (completed) iteration (\textit{it.
 time}) are also reported.
-The average of these values across the different values for $\Gamma$ are also provided for each instance set.
+The average of the reported values across the different values for $\Gamma$ are also given for each instance set.
+
+% Explanation of Benders' getting stuck in first iteration
+% Note also that when the Benders' algorithm is unable to find a feasible solution to a particular instance within the given time limit, this means that the algorithm was unable to solve the first iteration of the master problem.
+Hence this first iteration was not completed and is not counted in the reported \textit{it.} and \text{it.
+time} values.
+These instances are however recorded as having a solution time of 7200 seconds which is the reason for the disparity between the average iteration time and the average solution time in the $\Gamma=0$ row of the $j20$ set.
+
+Firstly and unsurprisingly, the results in Table \ref{table:benders_vs_compact} show that for both methods, the instances tend to increase in difficulty as $\Gamma$ increases.
+For the nominal problems, i.e.
+when $\Gamma=0$, the Benders' approach has a slight edge on the compact formulation, as we would expect given the relationship between the Benders' master problem formulation and the compact formulation.
+However, for all the non-zero values of $\Gamma$ across both instance sets, the compact formulation performs significantly stronger than the Benders' approach, solving a greater proportion of instances with dramatically reduced computation times.
 
@@
 		     \hline
-	\multirow{4}{*}{$j20$} & 0    & & 89.7     & 0.00     & 1            & 176.62        & 901.3        & & 89.4       & 1.92       & 941.9           \\
-			       & 5    & & 51.3     & 10.04    & 228          & 135.66        & 4153.6       & & 88.4       & 1.91       & 1061.3          \\
-			       & 10   & & 50.7     & 10.81    & 228          & 140.37        & 4230.4       & & 87.7       & 2.63       & 1138.4          \\
-			       & 15   & & 49.8     & 11.05    & 222          & 148.72        & 4287.1       & & 86.0       & 2.96       & 1279.8          \\
+	\multirow{4}{*}{$j20$} & 0    & & 89.9     & 0.00     & 1            & 171.08        & 885.5        & & 89.5       & 1.89       & 925.1           \\
+			       & 5    & & 50.9     & 10.21    & 220          & 133.32        & 4169.3       & & 88.6       & 1.87       & 1041.5          \\
+			       & 10   & & 50.4     & 10.99    & 225          & 138.49        & 4244.6       & & 87.9       & 2.58       & 1118.5          \\
+			       & 15   & & 49.5     & 11.21    & 216          & 146.26        & 4300.3       & & 86.3       & 2.91       & 1260.6          \\
 		     \cline{4-8} \cline{10-12}
-			       &      & & 60.4	   & 7.98     & 170	     & 150.34	     & 3393.1       & & 87.9       & 2.36       & 1105.4          \\
+			       &      & & 60.2	   & 8.10     & 165	     & 147.28	     & 3399.9       & & 88.1       & 2.31       & 1086.4          \\
 		     \hline \hline
@@
 }
-\caption{Comparison of the Benders' decomposition approach and the compact reformulation across the $j10$ and $j20$ instance sets and the different values of $\Gamma$.}
+\caption{Comparison of the Benders' decomposition approach and the compact reformulation across the $j10$ and $j20$ instance sets and for different values of $\Gamma$.}
 \label{table:benders_vs_compact}
@@
 
+Figures \ref{fig:j10_plots} and \ref{fig:j20_plots} show performance profiles and optimality gap plots for the compact formulation, as well as for both implementations of the Benders' decomposition solution approach (Algorithm \ref{alg:benders} and \cite{balouka2021robust}), across the $j10$ and $j20$ sets respectively.
+Note that only the results for the instances with non-zero values for $\Gamma$ were used to generate these plots, resulting in a total of 1608 $j10$ instances and 1662 $j20$ instances.
+
+Performance profiles (\cite{dolan2002benchmarking}) are plotted to compare competing solution approaches using their \textit{performance ratios}.
+The performance ratio of algorithm $a\in \mathcal{A}$ for instance $i\in \mathcal{I}$ is defined to be
+$$r_{ia}=\frac{t_{ia}}{\min_{a\in \mathcal{A}} t_{ia}},$$
+where $t_{ia}$ is the time required to solve instance $i$ using algorithm $a$.
+If method $a$ is unable to solve an instance $i$ within the 2-hour time limit, then $r_{ia}=R$, where $R\geq \max_{i,\,m}r_{im}$.
+The performance profiles in Figures \ref{fig:j10_performance_profile} and \ref{fig:j20_performance_profile} show the percentage of instances that have a performance ratio within a factor of $\tau$ of the best algorithm.
+These are plotted using a log scale for clarity.
+The y-intercepts of each method gives the percentage of instances for which that method found the optimal solution in the fastest time, whereas the right-most value gives the percentage of instances that were solved to optimality within the time limit.
+
+Figures \ref{fig:j10_gaps} and \ref{fig:j20_gaps} show the percentage of instances solved by each method to within a given optimality gap within the 2-hour time limit.
+These plots serve as a continuation of the performance profiles beyond just the instances that were solved to optimality.
+
+Looking at these plots, the improvements to the Benders' algorithm afforded by amending the master problem to use the formulation (\ref{eqn:master_first})-(\ref{eqn:master_last}) are clear.
+Across both the $j10$ and $j20$ instance sets, the \cite{balouka2021robust} Benders' implementation solves 55.1\% of instances to optimality and finds feasible solutions for a further 35.6\% of instances.
+Our Benders' implementation solves 65.0\% of instances and finds feasible solution solutions for a further 30.0\% of instances.
+
+We also see in these plots that the compact formulation completely dominates both Benders' methods.
+In particular across the two instances sets, the compact formulation finds a feasible solution to every instance, solving 93.1\% these to optimality.
+
+Comparing the solution times of the compact formulation with the stronger of the two Benders' implementations across the instances in $j10$ for which both methods found the optimal solution, the average solution time of the compact formulation (1.4s) was almost 200 times faster than for the Benders' approach (283.5s).
+And when both methods found the optimal solution for instances in the $j20$ set, the average solution time of the compact formulation (13.7s) was almost 100 times faster than for the Benders' approach (1304.5s).
+
 \begin{figure}[htb]
 	\makebox[\textwidth][c]{
-	\begin{subfigure}[t]{.65\textwidth}
+	\begin{subfigure}[t]{.6\textwidth}
 		\centering
 		\includegraphics[width=1.05\textwidth]{j10_performance_profile.pdf}
-		\caption{Performance profile of relative solution times.}\label{fig:j10_performance_profile}
+		\caption{Performance profiles showing percentage of instances with a performance ratio of within $\tau$. Plotted on a log scale.}\label{fig:j10_performance_profile}
 	\end{subfigure}%
-	\begin{subfigure}[t]{.65\textwidth}
+	\qquad
+	\begin{subfigure}[t]{.6\textwidth}
 		\centering
 		\includegraphics[width=1.05\textwidth]{j10_gaps.pdf}
-		\caption{Percentage of instances solved to within given gap of optimality within the two-hour time limit.}\label{fig:j10_gaps}
+		\caption{Percentage of instances solved to within given optimality gap within the two-hour time limit.}\label{fig:j10_gaps}
 	\end{subfigure}
 	}
-	\caption{Comparison of Benders' approach and compact reformulation over instances in the j10 set.}
+	\caption{Comparison of Benders' approaches and compact reformulation over instances in the j10 set.}
 	\label{fig:j10_plots}
@@
 	\makebox[\textwidth][c]{
-	\begin{subfigure}[t]{.65\textwidth}
+	\begin{subfigure}[t]{.6\textwidth}
 		\centering
 		\includegraphics[width=1.05\textwidth]{j20_performance_profile.pdf}
-		\caption{Performance profile of relative solution times.}\label{fig:j20_performance_profile}
+		\caption{Performance profile showing percentage of instances with a performance ratio of within $\tau$. Plotted on a log scale.}\label{fig:j20_performance_profile}
 	\end{subfigure}%
-	\begin{subfigure}[t]{.65\textwidth}
+	\qquad
+	\begin{subfigure}[t]{.6\textwidth}
 		\centering
 		\includegraphics[width=1.05\textwidth]{j20_gaps.pdf}
-		\caption{Percentage of instances solved to within given gap of optimality within the two-hour time limit.}\label{fig:j20_gaps}
+		\caption{Percentage of instances solved to within given optimality gap within the two-hour time limit.}\label{fig:j20_gaps}
 	\end{subfigure}
@@
 
+This work presented in this paper extends the compact mixed-integer programming formulation first introduced by \cite{bold2021compact} for application to the two-stage adjustable robust MRCPSP with uncertain activity durations.
+The computational performance of this formulation has been examined over a total of 3272 uncertain MRCPSP instances of varying size and difficulty, and compared against an improved version of the current state-of-the-art for solving this problem, based on a Benders' decomposition approach.
+The improved Benders' approach is the result of replacing the original master problem with a new formulation based on a simplified version of the compact formulation we present for the full problem.
+
+Results presented in Section \ref{section:experiments} show that the compact formulation we propose completely dominates the enhanced Benders' approach, solving over 43\% more instances to optimality, and doing so with dramatically reduced computation times.
+In addition to these strong computational improvements, the proposed compact formulation has the significant added benefit of being simpler to implement than the iterative Benders' approach.
+
+Despite these strong results, the instances used in these experiments contain only up to 20 activities.
+To enable the solving of larger scale instances, the development of heuristic solution approaches should be a primary focus of future research on this problem.
+
 \section*{Acknowledgements}

### 2022-03-11 paper/paper.tex
Complete first draft. Ready for Marc to review.

@@
 \bibliographystyle{apalike}
-\usepackage{bibentry}  % inline referencing
-\renewcommand{\inlinebibentry}[1]{\cite{#1} \bibentry{#1}}
 \usepackage{hyperref}
@@
 
-\nobibliography*  % tells bibentry to load bibliography database
-
 \begin{abstract}
@@
 
-	The proposed mixed-integer linear programming formulation is an extension of the formulation proposed by \cite{bold2021compact} (\bibentry{bold2021compact}) for the problem without multiple activity modes.
-Computational experiments show that this easy-to-implement formulation can be solved many times faster than the current state-of-the-art for solving this problem, solving over 40\% more instances to optimality over the same set of benchmarking instances.
+Computational experiments show that this easy-to-implement formulation is many times faster to solve than the current Benders' decomposition-based state-of-the-art solution approach for this problem, whilst also being able to solve over 40\% more instances to optimality over the same set of benchmarking instances.
 \end{abstract}
@@
 The multi-mode resource-constrained project scheduling problem (MRCPSP) is a generalisation of the widely-studied resource-constrained project scheduling problem (RCPSP) to include multiple processing modes for each activity.
-The inclusion of these modes is to model situations in which there is more than one way of executing project activities, with each option having its own duration and resource requirements.
-The MRCPSP consists selecting the processing modes and start times for a given set of activities, subject to a set of precedence constraints and limited resource availability, with the objective of minimising the overall project duration, known as the makespan.
+The inclusion of these modes allows the modelling of situations in which there is more than one way of executing project activities, with each option having its own duration and resource requirements.
+The MRCPSP consists of selecting the processing modes and start times for a given set of activities, subject to a set of precedence constraints and limited resource availability, with the objective of minimising the overall project duration, known as the makespan.
 
@@
 Following this, the actual activity durations are realised and a complete schedule is computed.
-The aim of the two-stage robust MRCPSP is to find a feasible first-stage solution (i.e.
+The aim of the two-stage adjustable robust MRCPSP is to find a feasible first-stage solution (i.e.
 mode selection and sequencing decisions) in order to minimise the realised worse-case makespan, as computed in the second-stage.
@@
 
-In recent years this two-stage robust optimisation approach has been applied to the RCPSP.
-First to use this approach were \cite{artigues2013robust}, who present an iterative scenario-relaxation algorithm for this problem with the objective of minimising the worst-case absolute regret.
-\cite{bruni2017adjustable} propose a Benders'-style decomposition approach for solving the robust RCPSP with the objective of minimising the worst-case project makespan.
-This work was extended in \cite{bruni2018computational}, which presents a computational study comparing an additional Benders' decomposition approach against a primal decomposition algorithm.
-Recently, \cite{bold2021compact} introduced a compact reformulation of the robust RCPSP and presented results which showed the superiority of that formulation over the iterative methods developed in those papers.
+A number of papers in recent years have applied this two-stage robust optimisation approach to the RCPSP.
+First to use this approach were \cite{artigues2013robust}, who presented an iterative scenario-relaxation algorithm for this problem with the objective of minimising the worst-case absolute regret.
+More recently, \cite{bruni2017adjustable} introduced a Benders'-style decomposition approach for solving the robust RCPSP with the objective of minimising the worst-case project makespan.
+This work was extended in \cite{bruni2018computational}, which presented a computational study comparing an additional Benders' decomposition approach against a primal decomposition algorithm.
+Most recently, \cite{bold2021compact} introduced a compact reformulation of the robust RCPSP and presented results which showed the superior computational performance of that formulation over the iterative decomposition-based methods developed in the two preceding papers.
 
-The application of the this two-stage robust optimisation approach for the MRCPSP is a more recently development.
-To the best of our knowledge, the only existing paper to consider this problem is \cite{balouka2021robust}, in which the Benders' decomposition approach used by \cite{bruni2017adjustable} for the robust RCPSP has been extended for application to the MRCPSP.
-Analogous to that work, the aim of this paper is to apply the compact reformulation first developed for the RCPSP by \cite{bold2021compact} to the MRCPSP.
-Specifically, we present a computational comparison of our compact reformulation approach and a strengthened version of the Benders' decomposition approach.
+The application of this two-stage robust optimisation approach for the MRCPSP is a very recent development.
+To the best of our knowledge, the only existing paper to consider this problem is \cite{balouka2021robust}, in which the Benders' decomposition approach introduced by \cite{bruni2017adjustable} for the robust RCPSP has been extended for application to the MRCPSP.
+Mirroring that extension, this paper adapts the compact formulation developed by \cite{bold2021compact} to the MRCPSP, with the aim of achieving similarly superior computational performance over the Benders' solution approach.
 
-\section{Problem description}
+Following a formal description of the two-stage robust MRCPSP in Section \ref{section:problem_description}, we outline the proposed compact formulation for this problem in Section \ref{section:compact_formulation} and present a strengthened version of the Benders' decomposition solution approach from \cite{balouka2021robust} in Section \ref{section:benders}.
+Results from a computational comparison of these two approaches are detailed in Section \ref{section:experiments}.
+
+\section{Problem description}\label{section:problem_description}
 
@@
 Each activity $i\in V$ has a set of available processing modes given by $M_i=\{1,\dots,|M_i|\}$.
-The nominal duration of activity $i$ when executed in mode $m_i\in M_i$ is given by $\bar{d}_{im_i}$, whilst its worst-case duration is given by $\bar{d}_{im_i}+\hat{d}_{im_i}$, where $\hat{d}_{im_i}$ is its maximum durational deviation.
-Each mode for activity $i$, $m_i\in M_i$, has an associated renewable resource requirement of $r_{im_ik}$ for each $k\in K$, where $K$ is the set of renewable resource types involved in the project.
+The nominal duration of activity $i$ when executed in mode $m\in M_i$ is given by $\bar{d}_{im}$, whilst its worst-case duration is given by $\bar{d}_{im}+\hat{d}_{im}$, where $\hat{d}_{im}$ is its maximum durational deviation.
+Each mode for activity $i$, $m\in M_i$, has an associated renewable resource requirement of $r_{imk}$ for each $k\in K$, where $K$ is the set of renewable resource types involved in the project.
 Each renewable resource $k\in K$ has an availability of $R_{k}$ at each time period in the project horizon.
-As well as renewable resource requirements, mode $m_i$ of activity $i$ also has a non-renewable resource requirement of $r^{'}_{im_ik}$ for each non-renewable resource type $k\in K^{'}$, with each non-renewable resource having an overall availability of $R^{'}_k$ for the entire project horizon.
+As well as renewable resource requirements, mode $m$ of activity $i$ also has a non-renewable resource requirement of $r^{'}_{imk}$ for each non-renewable resource type $k\in K^{'}$, with each non-renewable resource having an overall availability of $R^{'}_k$ for the entire project horizon.
 Additionally, the project is subject to a set of strict finish-to-start precedence constraints given by $E$, where $(i,j)\in E$ enforces that activity $i$ must finish before activity $j$ can begin.
@@
 	\centering
-	\includegraphics[scale=1]{mrcpsp_network.pdf}
+	\includegraphics[scale=0.9]{mrcpsp_network.pdf}
 	\caption{An example instance involving five non-dummy activities and a single renewable resource.
@@
 For a given choice of processing modes for each activity $\bm{m}=(m_1,\dots,m_n)$, we assume that the activity durations lie somewhere in a budgeted uncertainty set of the form
-$$\U_{\bm{m}}(\Gamma)=\Bigg\{\bm{d}\in\mathbb{R}_+^{|V|}:d_{im_i}=\bar{d}_{im_i}+\xi_i\hat{d}_{im_i},\,0\leq \xi_i \leq 1,\,\forall i\in V,\,\sum_{i\in V}\xi_i \leq \Gamma\Bigg\}.$$
+$$\U_{\bm{m}}(\Gamma)=\Bigg\{\bm{d}_{\bm{m}}\in\mathbb{R}_+^{|V|}:d_{im_i}=\bar{d}_{im_i}+\xi_i\hat{d}_{im_i},\,0\leq \xi_i \leq 1,\,\forall i\in V,\,\sum_{i\in V}\xi_i \leq \Gamma\Bigg\}.$$
 Introduced by \cite{bertsimas2004price}, the motivation of the budgeted uncertainty set is to control the pessimism of the solution by introducing a robustness parameter $\Gamma$ to limit the number of jobs that can simultaneously achieve their worst-case durations.
@@
 That is, a forbidden set is a collection of activities that cannot be executed in parallel only because of resource limitations.
-
 Applying the main representation theorem of \cite{bartusch1988scheduling}, for a particular choice of activity modes $\bm{m}$, a solution to the MRCPSP can be defined by a set of additional precedences $X_{\bm{m}}\subseteq V^2\setminus E$ such that the extended precedence network $G(V,E\cup X_{\bm{m}})$ is acyclic and contains no forbidden sets.
@@
 
-Hence the two-stage robust MRCPSP under budgeted uncertainty, which we refer to as the robust MRCPSP, can be written as
+The aim of the two-stage robust MRCPSP is therefore to select set of activity modes and a corresponding sufficient selection in order to minimise the worst-case project makespan.
+For the case where activity durations lie in a budgeted uncertainty set that we consider in this paper, this problem can be written as
 \begin{align}
@@
 where $\M\subseteq \mathbb{N}^n$ represents the set of all possible combinations of activity processing mode selections, and $\X_{\bm{m}}$ is the set of all possible sufficient selections for the choice of processing modes given by $\bm{m}$.
-This problem aims to selects activity modes and a corresponding sufficient selection in order to minimise the worst-case project makespan.
 
-\section{A compact formulation}
+An optimal solution to the instance shown in Figure \ref{fig:mrcpsp_network} is given by Figure \ref{fig:mrcpsp_network_solution}, where the mode choices are shown in bold, and the sufficient selection is given by $\{(3,2)\}$.
+The worst-case schedule for this optimal solution is shown in Figure \ref{fig:mrcpsp_schedule}, where activities 1 and 3 have been delayed to achieve a worst-case makespan of 15 (note that delaying a combination of activity 1 and any other activity results in the same worst-case makespan).
 
-We propose solving the robust MRCPSP using an extended version of the mixed-integer programming formulation developed by \cite{bold2021compact} for solving the robust RCPSP.
-This formulation combines the first and second-stage problems into a single compact formulation and was shown to significantly outperform the strongest iterative decomposition-based methods for the robust RCPSP.
+\begin{figure}[htb]
+	\makebox[\textwidth][c]{
+	\begin{subfigure}[t]{.5\textwidth}
+		\centering
+		\includegraphics[width=0.95\textwidth]{mrcpsp_solution.pdf}
+		\caption{Activity mode choices are highlighted in bold and sufficient selection is given by dashed arc $\{(3,2)\}$.}\label{fig:mrcpsp_network_solution}
+	\end{subfigure}%
+	\hspace{15mm}
+	\begin{subfigure}[t]{.6\textwidth}
+		\centering
+		\includegraphics[width=1.00\textwidth]{mrcpsp_schedule.pdf}
+		\caption{Worst-case schedule corresponding to optimal robust solution, where activities 1 and 3 have been delayed.}\label{fig:mrcpsp_schedule}
+	\end{subfigure}
+	}
+	\caption{Optimal solution to the example instance from Figure \ref{fig:mrcpsp_network}.}
+	\label{fig:mrcpsp_solution}
+\end{figure}
+
+\section{A compact formulation}\label{section:compact_formulation}
+
+We propose solving the robust MRCPSP using an extended version of the mixed-integer programming formulation developed by \cite{bold2021compact} for solving the two-stage adjustable robust RCPSP.
+This formulation combines the first and second-stage problems into a single compact formulation and was shown to significantly outperform the strongest iterative decomposition-based methods for that problem.
 
@@
 This augmented project network is formed of $\Gamma+1$ connected copies of the original project network, with each level being used to account for a delay to a single activity in the project.
-Having reformulated the adversarial subproblem as a longest-path problem, it can be dualised an inserted into a formulation for the first-stage problem, to arrive at a complete compact formulation for the full problem.
+Having reformulated the adversarial subproblem as a longest-path problem, strong duality can be employed to enable its insertion it into a formulation for the first-stage problem, and ultimately arrive at a complete compact formulation for the full problem.
 Given that the derivation of this formulation is more or less identical to the derivation of the corresponding formulation for the robust RCPSP, we omit it here and instead refer the reader to \cite{bold2021compact}.
@@
 Variables $x_{im}$ determine the processing mode for each activity, whilst resource flow variables $f_{ijk}$ track the amount of renewable resource $k$ that is transferred from activity $i$ to activity $j$ upon its completion.
-The $S_{i\gamma}$ variables are dual variables associated with the longest-path subproblem, which in this formulation these variables are used to track the start time of activities under the worst-case activity durations.
+The $S_{i\gamma}$ variables are dual variables associated with the longest-path subproblem, which in this formulation are used to track the start time of activities under the worst-case activity durations.
 Using these variables, the compact formulation for the robust MRCPSP can be written as
@@
 
-\section{A Benders' decomposition approach}
+\section{A Benders' decomposition approach}\label{section:benders}
 
 In this section we outline an alternative approach to solving the robust MRCPSP based on Benders' decomposition.
-Such a solution method for the robust MRCPSP was first presented in \cite{balouka2021robust}, and as mentioned previously, this itself was based on the approach to solve the robust RCPSP used in \cite{bruni2017adjustable}.
+A Benders'-type approach for solving the robust MRCPSP was first presented in \cite{balouka2021robust}, and as mentioned previously, this itself was based on the approach for solving the robust RCPSP used in \cite{bruni2017adjustable}.
 
-The main idea behind this approach is to decompose the full problem into a two smaller problems: 1.
+The main idea behind this approach is to decompose the full problem into its two stages: 1.
 a master problem that determines activity processing modes and resolves resource conflicts, and 2.
 a subproblem that takes the solution from the master problem and evaluates it by finding the worst-case makespan for the resulting network by solving a longest path problem.
-The solution to the master problem forms a lower bound to the optimal objective value of the original problem, whilst the solution to the subproblem provides an upper bound.
+Since the master problem does not account for all the problem uncertainty at once, its solution forms a lower bound to the optimal objective value of the original problem.
+Meanwhile, the solution to the subproblem is feasible to the original problem, and therefore provides an upper bound.
 These two problems are solved iteratively, with optimality cuts being added to the master problem each iteration based on the solution to the subproblem until the lower and upper bounds meet.
@@
 After initially implementing the Benders' algorithm as it is presented in \cite{balouka2021robust}, it was discovered that its performance could be strengthened by replacing the master problem formulation used in that implementation with a stripped-down version of the compact formulation (\ref{eqn:compact1})-(\ref{eqn:compact13}) in which the uncertainty is removed.
-In this section we therefore present our strengthened implementation of the Benders' approach for solving the robust MRCPSP.
-Note that a comparison of these two Benders' implementations is included in Section \ref{section:experiments}.
+In this section we present our strengthened implementation of the Benders' approach for solving the robust MRCPSP.
+A comparison of these two Benders' implementations is included in the results in Section \ref{section:experiments}.
 For the details of the original implementation of the Benders' decomposition algorithm for the robust MRCPSP, see \cite{balouka2021robust}.
@@
 
-The subproblem at iteration $t$ considers the worst-case makespan for the mode choices and sufficient selection from the master problem at iteration $t$, denoted by $\bm{m}^{*t}$ and $X_{\bm{m}^{*t}}$ respectively.
-This is done by solving an uncertain longest-path problem through the extended project network $T(E\cup X_{\bm{m}^{*t}})$, where activity durations lie in the budgeted uncertainty set $\U_{\bm{m}^{*t}}(\Gamma)$, which can be formulated as
+The subproblem at iteration $t$ computes the worst-case makespan for the mode choices and sufficient selection from the master problem at iteration $t$, denoted by $\bm{m}^{*t}$ and $X_{\bm{m}^{*t}}$ respectively.
+This is done by solving a longest-path problem through the extended project network $T(E\cup X_{\bm{m}^{*t}})$, in which the activity durations (i.e.
+arc lengths) can be chosen from the budgeted uncertainty set $\U_{\bm{m}^{*t}}(\Gamma)$.
+This problem can be formulated as
 \begin{align}
@@
 
-Variables $\alpha_{ij}$ define the longest path through the network.
-This path, which is denote as $\pi^{*t}=\{(i,j),\,i,j\in V:\alpha_{ij}=1\}$ has length $V^{*t}$.
+Variables $\alpha_{ij}$ define the longest path through the network, which we denote as $\pi^{*t}=\{(i,j),\,i,j\in V:\alpha_{ij}=1\}$, and has length $V^{*t}$.
 Variables $\xi_i$ are the activity delay variables, whilst $w_{ij}$ are used to linearise the formulation.
@@
 
-Given the current best lower bound for the original problem $LB$, the mode selection from the master problem $\bm{m}^{*t}$, and the optimal objective value of the subproblem $V^{*t}$, the cut at iteration $t$ can be written as
+Given the current best lower bound for the original problem, $LB$, the mode selection from the master problem, $\bm{m}^{*t}$, and the optimal objective value of the subproblem, $V^{*t}$, the cut at iteration $t$ can be written as
 \begin{align}
 	& \eta \geq (V^{*t} - LB)\cdot\sum_{(i,j)\in \pi^{*t}}\bigg(1/3 (y_{ij} + x_{i,m_i^{*t}} + x_{j,m_j^{*t}}) - (3 - y_{ij} - x_{i,m_i^{*t}} - x_{j,m_j^{*t}})\bigg) & \nonumber \\
-	& \hspace{8.1cm} - (V^{*t} - LB)\cdot(|\pi^{*t}|-1) + LB,\label{eqn:benders_cut}
+	& \hspace{8.1cm} - (V^{*t} - LB)\cdot(|\pi^{*t}|-1) + LB.\label{eqn:benders_cut}
 \end{align}
@@
 
-\subsection{Example}
-
-Here, to demonstrate its implementation, we use the Benders' decomposition approach to solve the example shown in Figure \ref{fig:mrcpsp_network} with $R_1=4$ and $\Gamma=2$.
-An outline of the implementation of the Benders' approach is presented in Algorithm \ref{alg:benders}.
-
-The algorithm solves this instance in 6 iterations, and the solution information from the master and subproblem at each of these iterations is shown in Table \ref{table:benders_example}.
-Note that in the final iteration, the algorithm can terminate after solving the master problem and updating the $UB$ since $LB=UB$, i.e.
-there is no need to solve the subproblem in the final iteration.
+An overview of the implementation of the Benders' solution approach outlined in this section is presented in Algorithm \ref{alg:benders}.
 
@@
 
+\subsection{Example}
+
+To demonstrate its implementation, we use the Benders' decomposition approach to solve the example shown in Figure \ref{fig:mrcpsp_network} with $R_1=4$ and $\Gamma=2$.
+The algorithm solves this instance in 6 iterations, and the solution information from the master and subproblem at each of these iterations is shown in Table \ref{table:benders_example}.
+
+The first iteration of the master problem solves the nominal instance, assuming no delays to the activity durations.
+This solution opts to schedule activities 2 and 3 in parallel, which can be achieved by setting $m_2=2$, and the sufficient selection for is given by $\{(4,5)\}$.
+This first solution is evaluated in the subproblem to have a worst-case makespan of 16.
+In the second iteration, having added the first optimality cut, the master problem finds the solution shown in Figure \ref{fig:mrcpsp_network_solution}.
+By calculating the schedule shown in Figure \ref{fig:mrcpsp_schedule}, the subproblem evaluates the objective value of this solution to be 15.
+Although this solution is optimal, the algorithm requires a further four iterations to prove that this is the case.
+Note that there is no need to solve the subproblem in the final iteration since after solving the master problem and updating $UB$, we have that $LB=UB$.
 \begin{table}[h]
@@
 \end{tabular}
-\caption{Solution information at each of the six iterations of the Benders' decomposition approach required to solve example instance in Figure \ref{fig:mrcpsp_network}.}
+\caption{Solution information at each of the six iterations of the Benders' algorithm required to solve example instance in Figure \ref{fig:mrcpsp_network}.}
 \label{table:benders_example}
@@
 
-The instances used in this computational study have been created using the deterministic $j10$, and $j20$ MRCPSP instances from the PSPLIB (\cite{kolisch1997psplib}, \url{https://www.om-db.wi.tum.de/psplib/}).
+The instances used in this computational study have been created from the deterministic $j10$, and $j20$ MRCPSP instances from the PSPLIB (\cite{kolisch1997psplib}, \url{https://www.om-db.wi.tum.de/psplib/}).
 The $j10$ set contains a total of 536 instances each involving 10 activities, and the $j20$ set contains a total of 554 instances each involving 20 activities.
@@
 The values in these tables have been computed only over the instances for which an optimal solution was found for all the values of $\Gamma$ (i.e.
-all 536 instances in the $j10$ set and 477 instances of the $j20$ set).
-As expected, as $\Gamma$ increases the solution cost increases in a concave manner.
+all 536 instances in the $j10$ set, and 477 instances of the $j20$ set).
+As we would expect, the solution cost increases in a concave manner as $\Gamma$ increases.
 
@@
 
-Firstly and unsurprisingly, the results in Table \ref{table:benders_vs_compact} show that for both methods, the instances tend to increase in difficulty as $\Gamma$ increases.
-For the nominal problems, i.e.
-when $\Gamma=0$, the Benders' approach has a slight edge on the compact formulation, as we would expect given the relationship between the Benders' master problem formulation and the compact formulation.
-However, for all the non-zero values of $\Gamma$ across both instance sets, the compact formulation performs significantly stronger than the Benders' approach, solving a greater proportion of instances with dramatically reduced computation times.
+Firstly, the results in Table \ref{table:benders_vs_compact} show that for both methods, the instances tend to increase in difficulty as $\Gamma$ increases.
+We can also see that for the nominal problems, i.e.
+when $\Gamma=0$, the Benders' approach has a slight edge on the compact formulation.
+This is what we would expect given the relationship between the Benders' master problem formulation and the compact formulation.
+However, for all the non-zero values of $\Gamma$ across both instance sets, the compact formulation performs significantly better than the Benders' approach, solving a greater proportion of instances with dramatically reduced computation times.
 
@@
 
-Figures \ref{fig:j10_plots} and \ref{fig:j20_plots} show performance profiles and optimality gap plots for the compact formulation, as well as for both implementations of the Benders' decomposition solution approach (Algorithm \ref{alg:benders} and \cite{balouka2021robust}), across the $j10$ and $j20$ sets respectively.
-Note that only the results for the instances with non-zero values for $\Gamma$ were used to generate these plots, resulting in a total of 1608 $j10$ instances and 1662 $j20$ instances.
+Figures \ref{fig:j10_plots} and \ref{fig:j20_plots} show performance profiles and optimality gap plots for the compact formulation and both implementations of the Benders' decomposition solution approach (Algorithm \ref{alg:benders} and \cite{balouka2021robust}), across the $j10$ and $j20$ sets respectively.
+Note that only the results for the instances with non-zero values for $\Gamma$ were used to generate these plots, resulting in a total of 1608 $j10$ instances and 1664 $j20$ instances.
 
-Performance profiles (\cite{dolan2002benchmarking}) are plotted to compare competing solution approaches using their \textit{performance ratios}.
+Performance profiles (\cite{dolan2002benchmarking}) provide a visual comparison of competing solution approaches using their \textit{performance ratios}.
 The performance ratio of algorithm $a\in \mathcal{A}$ for instance $i\in \mathcal{I}$ is defined to be
@@
 These are plotted using a log scale for clarity.
-The y-intercepts of each method gives the percentage of instances for which that method found the optimal solution in the fastest time, whereas the right-most value gives the percentage of instances that were solved to optimality within the time limit.
+The y-intercepts of each method give the percentage of instances for which that method found the optimal solution in the fastest time, whereas the right-most value gives the overall percentage of instances that were solved to optimality within the time limit by that method.
+
+These performance profiles show that whenever an instance was able to be solved to optimality by either solution approach, the compact formulation was always the faster method.
+If we specifically compare the solution times of the compact formulation with the stronger of the two Benders' implementations, across the instances in $j10$ for which both methods found the optimal solution, the average solution time of the compact formulation (1.4s) was almost 200 times faster than for the Benders' approach (283.5s).
+And when both methods found the optimal solution for instances in the $j20$ set, the average solution time of the compact formulation (13.7s) was almost 100 times faster than for the Benders' approach (1304.5s).
 
@@
 These plots serve as a continuation of the performance profiles beyond just the instances that were solved to optimality.
+Summarising the data from these plots across both the $j10$ and $j20$ instance sets, the \cite{balouka2021robust} Benders' implementation solves 55.1\% of instances to optimality and finds feasible solutions for a further 35.6\% of instances.
+Our Benders' implementation solves 65.0\% of instances and finds feasible solution solutions for a further 30.0\% of instances.
+The compact formulation on the other hand finds a feasible solution to every instance, solving 93.1\% these to optimality.
 
-Looking at these plots, the improvements to the Benders' algorithm afforded by amending the master problem to use the formulation (\ref{eqn:master_first})-(\ref{eqn:master_last}) are clear.
-Across both the $j10$ and $j20$ instance sets, the \cite{balouka2021robust} Benders' implementation solves 55.1\% of instances to optimality and finds feasible solutions for a further 35.6\% of instances.
-Our Benders' implementation solves 65.0\% of instances and finds feasible solution solutions for a further 30.0\% of instances.
-
-We also see in these plots that the compact formulation completely dominates both Benders' methods.
-In particular across the two instances sets, the compact formulation finds a feasible solution to every instance, solving 93.1\% these to optimality.
-
-Comparing the solution times of the compact formulation with the stronger of the two Benders' implementations across the instances in $j10$ for which both methods found the optimal solution, the average solution time of the compact formulation (1.4s) was almost 200 times faster than for the Benders' approach (283.5s).
-And when both methods found the optimal solution for instances in the $j20$ set, the average solution time of the compact formulation (13.7s) was almost 100 times faster than for the Benders' approach (1304.5s).
-
-\begin{figure}[htb]
+The results presented here show the clear improvements to the Benders' algorithm afforded by amending the master problem to use the formulation (\ref{eqn:master_first})-(\ref{eqn:master_last}).
+More significantly however, these results demonstrate the complete dominance of the compact formulation over both Benders' methods.
+\begin{figure}[h]
 	\makebox[\textwidth][c]{
@@
 	}
-	\caption{Comparison of Benders' approaches and compact reformulation over instances in the j10 set.}
+	\caption{Comparison of the Benders' approaches and compact formulation over instances in the $j10$ set.}
 	\label{fig:j10_plots}
@@
 
-\begin{figure}[htb]
+\begin{figure}[h]
 	\makebox[\textwidth][c]{
@@
 	}
-	\caption{Comparison of Benders' approach and compact reformulation over instances in the j20 set.}
+	\caption{Comparison of the Benders' approaches and compact formulation over instances in the $j20$ set.}
 	\label{fig:j20_plots}
@@
 
-This work presented in this paper extends the compact mixed-integer programming formulation first introduced by \cite{bold2021compact} for application to the two-stage adjustable robust MRCPSP with uncertain activity durations.
-The computational performance of this formulation has been examined over a total of 3272 uncertain MRCPSP instances of varying size and difficulty, and compared against an improved version of the current state-of-the-art for solving this problem, based on a Benders' decomposition approach.
+The work presented in this paper extends the compact mixed-integer programming formulation first introduced by \cite{bold2021compact}, for application to the two-stage adjustable robust MRCPSP with uncertain activity durations.
+The computational performance of this formulation has been examined over a total of 3270 uncertain MRCPSP instances of varying size and difficulty, and compared against an improved version of the current state-of-the-art for solving this problem, based on a Benders' decomposition approach.
 The improved Benders' approach is the result of replacing the original master problem with a new formulation based on a simplified version of the compact formulation we present for the full problem.
 
-Results presented in Section \ref{section:experiments} show that the compact formulation we propose completely dominates the enhanced Benders' approach, solving over 43\% more instances to optimality, and doing so with dramatically reduced computation times.
+Results presented in Section \ref{section:experiments} show that the compact formulation completely dominates the enhanced Benders' approach, solving over 43\% more instances to optimality, and doing so with dramatically reduced computation times.
 In addition to these strong computational improvements, the proposed compact formulation has the significant added benefit of being simpler to implement than the iterative Benders' approach.

### 2022-03-11 code/solve_instance.py
Combining toy_example_solve.py into solve_instance.py.

@@ -1,14 +1,16 @@
+import mrcpsp
 from benders import Benders
 from compact_reformulation import CompactRefomulation
 
 
-def solve(instance, solve_method, Gamma, time_limit):
+def solve(nominal_data_file, solve_method, Gamma, time_limit, max_durational_deviations=None, print_log=False):
     """
-    Solves given uncertain MRCPSP instance using solve_method ('compact_reformulation' or 'benders') and prints solve
-    log to terminal. Used to test code for specific instances.
+    Solve instance given in nominal_data_file with Gamma using specified solve method ('compact_reformulation' or
+    'benders'). Max durational deviations can be explicitly provided for each activity in the input, but if not provided
+    they are set to floor(0.7 * nominal values). This function is used to test code for specific instances.
 
-    :param instance: Uncertain MRCPSP instance.
-    :type instance: Instance
+    :param nominal_data_file: Absolute file path to deterministic MRCPSP instance from which to create robust instance.
+    :type nominal_data_file: str
     :param solve_method: Solution method to use, i.e. 'compact_reformulation' or 'benders'.
     :type solve_method: str
     :param Gamma: Robustness parameter, i.e. max number of jobs that can achieve worst-case durations simultaneously.
@@ -16,25 +18,61 @@ def solve(instance, solve_method, Gamma, time_limit):
     :type Gamma: int
     :param time_limit: Time limit (in seconds) to give to chosen solution method.
     :type time_limit: int
+    :param max_durational_deviations: Optional parameter. Dictionary with max durational deviation for each mode of each
+        activity.
+    :type max_durational_deviations: dict
+    :param print_log: Indicates whether or not to print solve log to terminal. Defaults to False.
+    :type print_log: bool
     """
+    # load instance
+    instance = mrcpsp.load_nominal_mrcpsp(nominal_data_file)
+    if max_durational_deviations:
+        instance.set_dbar_explicitly(max_durational_deviations)
+    else:
+        instance.set_dbar_uncertainty_level(0.7)
+
     # check solve_method is recognised
     if solve_method not in ['compact_reformulation', 'benders']:
         raise Exception("{} is not a recognised solve method. Available methods are 'compact_reformulation' and "
[…]

### 2022-03-11 code/experiment.py
Removing misc_functions.py.

@@ -4,7 +4,6 @@ import argparse
 from mrcpsp import load_nominal_mrcpsp
 from benders import Benders
 from compact_reformulation import CompactRefomulation
-from misc_functions import is_valid_directory
 
 
 def create_instances(instance_dir, uncertainty_level):
@@ -105,6 +104,16 @@ def run_experiment(instances, Gamma, solve_method, time_limit):
                 f.write("{}\t{}\t{:.3f}\t{:.3f}\t{:.3f}\t{:.3f}\n".format(instance.name, Gamma, compact_sol['objval'],
                                                                           compact_sol['objbound'], gap,
                                                                           compact_sol['runtime']))
+
+
+def is_valid_directory(arg):
+    """
+    Checks directory path given as input argument exists. Throws an error if it does not.
+    """
+    if os.path.isdir(arg):
+        return arg
+    else:
+        raise argparse.ArgumentTypeError("{} is not a valid directory.".format(arg))
 
 
 if __name__ == '__main__':

### 2022-03-11 code/parallel_experiment.py

@@ -7,7 +7,6 @@ from functools import partial
 from mrcpsp import load_nominal_mrcpsp
 from benders import Benders
 from compact_reformulation import CompactRefomulation
-from misc_functions import is_valid_directory
 
 
 def create_instances(instance_dir, uncertainty_level):
@@ -172,6 +171,16 @@ def reorder_results(results_file):
         f.write(first_line + '\n')
         for instance in instance_names:
             f.write(data[instance] + '\n')
+
+
+def is_valid_directory(arg):
+    """
+    Checks directory path given as input argument exists. Throws an error if it does not.
+    """
+    if os.path.isdir(arg):
+        return arg
+    else:
+        raise argparse.ArgumentTypeError("{} is not a valid directory.".format(arg))
 
 
 if __name__ == '__main__':

### 2022-03-11 code/toy_example_solve.py

@@ -1,60 +1,0 @@
-import mrcpsp
-from compact_reformulation import CompactRefomulation
-from benders import Benders
-
-
-def solve_toy_example(nominal_data_file, Gamma, max_durational_deviations=None):
-    """
-    Solves toy example given in nominal_data_file for given Gamma. Max durational deviations can be explicitly provided
-    for each activity in the input, but if not provided they are set to 0.5 * nominal values.
-    """
-    # load instance
-    instance = mrcpsp.load_nominal_mrcpsp(nominal_data_file)
-    if max_durational_deviations:
-        instance.set_dbar_explicitly(max_durational_deviations)
-    else:
-        instance.set_dbar_uncertainty_level(0.5)
-
-    print(f"\nSolving toy example given in {nominal_data_file}...")
-    print("\nJob data [r, d, d_bar]:")
-    print("-------------------------")
-    for i in instance.V:
-        print("job {}:".format(i),
-              [[instance.jobs[i].r[m][0], instance.jobs[i].d[m], instance.jobs[i].d_bar[m]] for m in
-               instance.jobs[i].M])
-
-    print("\nCompact reformulation")
-    print("---------------------")
-    compact_sol = CompactRefomulation(instance, Gamma, time_limit=20).solve(print_log=False)
-    print("objval:", compact_sol['objval'])
-    print("runtime:", compact_sol['runtime'])
-    print("modes:", compact_sol['modes'])
-    print("network:", compact_sol['network'])
-    print("resource flows:", compact_sol['flows'])
-
-    print("\nBenders'")
-    print("--------")
-    benders_sol = Benders(instance, Gamma, time_limit=20).solve(print_log=True)
-    print("objval:", benders_sol['objval'])
-    print("runtime:", benders_sol['runtime'])
-    print("n_iterations:", benders_sol['n_iterations'])
-    print("modes:", benders_sol['modes'])
-    print("network:", compact_sol['network'])
-    print("resource flows:", benders_sol['flows'])
-
-
-if __name__ == "__main__":
-
-    # toy example from Balouka & Cohen (2021)
-    balouka_toy_example = '/home/boldm1/OneDrive/robust-mrcpsp/code/instances/mrcpsp_toy_example_data.txt'
[…]

### 2022-03-11 paper/paper.tex
Fixing some warning messages in tex file.
small edits

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newcommand{\marc}[1]{\textcolor{blue}{#1}}
 
@@
 
-The aim of the two-stage robust MRCPSP is therefore to select set of activity modes and a corresponding sufficient selection in order to minimise the worst-case project makespan.
+The aim of the two-stage robust MRCPSP is therefore to 
+% select set of 
+\marc{determine}
+activity modes and a corresponding sufficient selection in order to minimise the worst-case project makespan.
 For the case where activity durations lie in a budgeted uncertainty set that we consider in this paper, this problem can be written as
 \begin{align}
-	& \min_{\bm{m}\in \M,\,X_{\bm{m}}\in \X_{\bm{m}}} \max_{\bm{d}\in\U_{\bm{m}}(\Gamma)} \min S_{n+1}\label{eqn:robust_mrcpsp_1}\\
-	& S_0 = 0\\
+	\min_{\bm{m}\in \M,\,X_{\bm{m}}\in \X_{\bm{m}}} \max_{\bm{d}\in\U_{\bm{m}}(\Gamma)}\ & \min S_{n+1}\label{eqn:robust_mrcpsp_1}\\
+	\text{s.t. } & S_0 = 0\\
 	& S_j - S_i \geq d_{im_i} \qquad \forall (i,j)\in E\cup X_{\bm{m}}\\
@@
 This formulation is constructed by recasting the adversarial subproblem of determining the worst-case activity durations for a given set of mode choices and sufficient selection as a longest path problem on an augmented project network.
-This augmented project network is formed of $\Gamma+1$ connected copies of the original project network, with each level being used to account for a delay to a single activity in the project.
+This
+% augmented project network 
+\marc{directed acyclic graph}
+is formed of $\Gamma+1$ connected copies of the original project network, with each level being used to account for a delay to a single activity in the project \marc{(a similar construction has been applied in \cite{bendotti2019anchor})}.
 Having reformulated the adversarial subproblem as a longest-path problem, strong duality can be employed to enable its insertion it into a formulation for the first-stage problem, and ultimately arrive at a complete compact formulation for the full problem.
@@
 Hence, to remove the uncertainty from the model, (\ref{eqn:compact1})-(\ref{eqn:compact13}) is modified by considering only a single level of the augmented project network.
-Using this formulation, the master problem can be written as
+Using this formulation, the master problem \marc{in iteration $t$} can be written as
 \begin{align}
@@
 	\State \textbf{Solve master problem (\ref{eqn:master_first})-(\ref{eqn:master_last})} 
-	\State \hspace{\algorithmicindent} Get objective value $\eta^{*t}$, processing modes $\bm{m^{*t}}$ and precedences $\bm{y}_{\bm{m}^{*t}}$
+	\State \hspace{\algorithmicindent} Get objective value $\eta^{*t}$, processing modes $\bm{m}^{*t}$ and precedences $\bm{y}_{\bm{m}^{*t}}$
 	\State \hspace{\algorithmicindent} If $\eta^{*t}>LB$, update $LB\leftarrow \eta^{*t}$ 
@@
 	\hline \hline
-	$t$ & $LB$ & $UB$ & \eta^{*t} & $V^{*t}$ & \multicolumn{1}{c}{$\bm{m}^{*t}$}     & \multicolumn{1}{c}{$\pi^{*t}$}  \\
+	$t$ & $LB$ & $UB$ & $\eta^{*t}$ & $V^{*t}$ & \multicolumn{1}{c}{$\bm{m}^{*t}$}     & \multicolumn{1}{c}{$\pi^{*t}$}  \\
 	\hline
@@
 \caption{Solution information at each of the six iterations of the Benders' algorithm required to solve example instance in Figure \ref{fig:mrcpsp_network}.}
-\label{table:benders_example}
+\label{table:benders_example}}
 \end{table}
@@
 	\caption{Average optimal objective values across instances in the $j10$ and $j20$ instance sets for different values of $\Gamma$.}
-	\label{table:objvals}
+	\label{table:objvals}}
 \end{table}
@@
 
-Performance profiles (\cite{dolan2002benchmarking}) provide a visual comparison of competing solution approaches using their \textit{performance ratios}.
-The performance ratio of algorithm $a\in \mathcal{A}$ for instance $i\in \mathcal{I}$ is defined to be
-$$r_{ia}=\frac{t_{ia}}{\min_{a\in \mathcal{A}} t_{ia}},$$
-where $t_{ia}$ is the time required to solve instance $i$ using algorithm $a$.
-If method $a$ is unable to solve an instance $i$ within the 2-hour time limit, then $r_{ia}=R$, where $R\geq \max_{i,\,m}r_{im}$.
-The performance profiles in Figures \ref{fig:j10_performance_profile} and \ref{fig:j20_performance_profile} show the percentage of instances that have a performance ratio within a factor of $\tau$ of the best algorithm.
-These are plotted using a log scale for clarity.
-The y-intercepts of each method give the percentage of instances for which that method found the optimal solution in the fastest time, whereas the right-most value gives the overall percentage of instances that were solved to optimality within the time limit by that method.
-
-These performance profiles show that whenever an instance was able to be solved to optimality by either solution approach, the compact formulation was always the faster method.
-If we specifically compare the solution times of the compact formulation with the stronger of the two Benders' implementations, across the instances in $j10$ for which both methods found the optimal solution, the average solution time of the compact formulation (1.4s) was almost 200 times faster than for the Benders' approach (283.5s).
-And when both methods found the optimal solution for instances in the $j20$ set, the average solution time of the compact formulation (13.7s) was almost 100 times faster than for the Benders' approach (1304.5s).
-
-Figures \ref{fig:j10_gaps} and \ref{fig:j20_gaps} show the percentage of instances solved by each method to within a given optimality gap within the 2-hour time limit.
-These plots serve as a continuation of the performance profiles beyond just the instances that were solved to optimality.
-Summarising the data from these plots across both the $j10$ and $j20$ instance sets, the \cite{balouka2021robust} Benders' implementation solves 55.1\% of instances to optimality and finds feasible solutions for a further 35.6\% of instances.
-Our Benders' implementation solves 65.0\% of instances and finds feasible solution solutions for a further 30.0\% of instances.
-The compact formulation on the other hand finds a feasible solution to every instance, solving 93.1\% these to optimality.
-
-The results presented here show the clear improvements to the Benders' algorithm afforded by amending the master problem to use the formulation (\ref{eqn:master_first})-(\ref{eqn:master_last}).
-More significantly however, these results demonstrate the complete dominance of the compact formulation over both Benders' methods.
-\begin{figure}[h]
+\begin{figure}[htb]
 	\makebox[\textwidth][c]{
@@
 
-\begin{figure}[h]
+\begin{figure}[htb]
 	\makebox[\textwidth][c]{
@@
 \end{figure}
+
+Performance profiles (\cite{dolan2002benchmarking}) provide a visual comparison of competing solution approaches using their \textit{performance ratios}.
+The performance ratio of algorithm $a\in \mathcal{A}$ for instance $i\in \mathcal{I}$ is defined to be
+$$r_{ia}=\frac{t_{ia}}{\min_{a\in \mathcal{A}} t_{ia}},$$
+where $t_{ia}$ is the time required to solve instance $i$ using algorithm $a$.
+If method $a$ is unable to solve an instance $i$ within the 2-hour time limit, then $r_{ia}=R$, where $R\geq \max_{i,\,m}r_{im}$.
+The performance profiles in Figures \ref{fig:j10_performance_profile} and \ref{fig:j20_performance_profile} show the percentage of instances that have a performance ratio within a factor of $\tau$ of the best algorithm.
+These are plotted using a log scale for clarity.
+The y-intercepts of each method give the percentage of instances for which that method found the optimal solution in the fastest time, whereas the right-most value gives the overall percentage of instances that were solved to optimality within the time limit by that method.
+
+These performance profiles show that whenever an instance was able to be solved to optimality by either solution approach, the compact formulation was always the faster method.
+If we specifically compare the solution times of the compact formulation with the stronger of the two Benders' implementations, across the instances in $j10$ for which both methods found the optimal solution, the average solution time of the compact formulation (1.4s) was almost 200 times faster than for the Benders' approach (283.5s).
+% And 
+\marc{When} both methods found the optimal solution for instances in the $j20$ set, the average solution time of the compact formulation (13.7s) was almost 100 times faster than for the Benders' approach (1304.5s).
+
+Figures \ref{fig:j10_gaps} and \ref{fig:j20_gaps} show the percentage of instances solved by each method to within a given optimality gap within the 2-hour time limit.
+These plots serve as a continuation of the performance profiles beyond just the instances that were solved to optimality.
+Summarising the data from these plots across both the $j10$ and $j20$ instance sets, the \cite{balouka2021robust} Benders' implementation solves 55.1\% of instances to optimality and finds feasible solutions for a further 35.6\% of instances.
+Our Benders' implementation solves 65.0\% of instances and finds feasible solution solutions for a further 30.0\% of instances.
+The compact formulation on the other hand finds a feasible solution to every instance, solving 93.1\% these to optimality.
+
+The results presented here show the clear improvements to the Benders' algorithm afforded by amending the master problem to use the formulation (\ref{eqn:master_first})-(\ref{eqn:master_last}).
+More significantly however, these results demonstrate the complete dominance of the compact formulation over both Benders' methods.
 

### 2022-03-13 paper/paper.tex
Responding to Marc's comments.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newcommand{\marc}[1]{\textcolor{blue}{#1}}
 
@@
 	\includegraphics[scale=0.9]{mrcpsp_network.pdf}
-	\caption{An example instance involving five non-dummy activities and a single renewable resource.
+	\caption{An example instance involving five non-dummy activities and a single renewable resource with availability $R_1=4$.
 Activities 2 and 5 each have two available processing modes, whilst the other activities have only a single mode.}
@@
 
-The aim of the two-stage robust MRCPSP is therefore to 
-% select set of 
-\marc{determine}
-activity modes and a corresponding sufficient selection in order to minimise the worst-case project makespan.
+The aim of the two-stage robust MRCPSP is therefore to determine activity modes and a corresponding sufficient selection in order to minimise the worst-case project makespan.
 For the case where activity durations lie in a budgeted uncertainty set that we consider in this paper, this problem can be written as
@@
 This formulation is constructed by recasting the adversarial subproblem of determining the worst-case activity durations for a given set of mode choices and sufficient selection as a longest path problem on an augmented project network.
-This
-% augmented project network 
-\marc{directed acyclic graph}
-is formed of $\Gamma+1$ connected copies of the original project network, with each level being used to account for a delay to a single activity in the project \marc{(a similar construction has been applied in \cite{bendotti2019anchor})}.
+This directed acyclic graph is formed of $\Gamma+1$ connected copies of the original project network, with each level being used to account for a delay to a single activity in the project (a similar construction has also been applied in \cite{bendotti2019anchor}).
 Having reformulated the adversarial subproblem as a longest-path problem, strong duality can be employed to enable its insertion it into a formulation for the first-stage problem, and ultimately arrive at a complete compact formulation for the full problem.
@@
 Hence, to remove the uncertainty from the model, (\ref{eqn:compact1})-(\ref{eqn:compact13}) is modified by considering only a single level of the augmented project network.
-Using this formulation, the master problem \marc{in iteration $t$} can be written as
+Using this formulation, the master problem at iteration $t$ can be written as
 \begin{align}
@@
 	     & S_{j} - S_{i} \geq \bar{d}_{im}x_{im} - N(1-y_{ij}) \qquad \forall (i,j)\in V^2,\,\forall m\in M_i\label{eqn:master3}\\
-	     & \eta \geq \lambda(x, y, M^{*\ell}) \qquad \forall \ell=1,\dots,t-1\label{eqn:master_cuts}\\
+	     & \eta \geq \lambda(\bm{x}, \bm{y}, V^{*\ell}) \qquad \forall \ell=1,\dots,t-1\label{eqn:master_cuts}\\
 	     & \textnormal{(\ref{eqn:compact5})-(\ref{eqn:compact_nonrenewable}),\,(\ref{eqn:compact11})-(\ref{eqn:compact13})}\\
@@
 The instances used in this computational study have been created from the deterministic $j10$, and $j20$ MRCPSP instances from the PSPLIB (\cite{kolisch1997psplib}, \url{https://www.om-db.wi.tum.de/psplib/}).
+Note that these are the same instance sets as used for the experiments in \cite{balouka2021robust}.
 The $j10$ set contains a total of 536 instances each involving 10 activities, and the $j20$ set contains a total of 554 instances each involving 20 activities.
@@
 If we specifically compare the solution times of the compact formulation with the stronger of the two Benders' implementations, across the instances in $j10$ for which both methods found the optimal solution, the average solution time of the compact formulation (1.4s) was almost 200 times faster than for the Benders' approach (283.5s).
-% And 
-\marc{When} both methods found the optimal solution for instances in the $j20$ set, the average solution time of the compact formulation (13.7s) was almost 100 times faster than for the Benders' approach (1304.5s).
+When both methods found the optimal solution for instances in the $j20$ set, the average solution time of the compact formulation (13.7s) was almost 100 times faster than for the Benders' approach (1304.5s).
 
